### 2017-05-25 EmptyHomesV1.Rmd
Remove DAT
Changed the way maps are plotted so the DAT dataframe is no longer needed. It also allows for all data on a region to be plotted even when that data is missing

---
title: "Empty Homes V1"
author: "Jonathan Bourne"
date: "10 February 2017"
output: html_document
---

It would have been sensible to separate council and private housing

PCLB0 is second homes, PCLC0 is long term empty, PCLC100 is empty and unfurnished up to six months and PCLD100 is undergoing major repairs up to 12 months.

#Telegraph article on pensioners requireing mortgages
http://www.telegraph.co.uk/personal-banking/mortgages/pensioner-mortgages-boom-britains-broken-housing-market/

Mean household income 31920
https://www.ons.gov.uk/peoplepopulationandcommunity/personalandhouseholdfinances/incomeandwealth/datasets/nowcastinghouseholdincomeintheuk

```{r setup}
setwd( "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode")
source("Setup.R")

setwd(Functions)
list.files() %>% map(source)
```

#VOA data

Council tax bands from VOA
https://www.gov.uk/government/statistics/council-tax-stock-of-properties-2016

```{r PopandHomes}
setwd(DataFolder)

CTtaxstock<- read_csv("Table_CTSOP1.1_2016.csv")

pop <- read_csv("SAPE2.csv")

EW <- inner_join(select(CTtaxstock, ECODE, ALL_PROPERTIES),select(pop, `Area Codes`, `All Ages`),
          by=c("ECODE" = "Area Codes")) %>% setNames(c("ECODE", "Homes", "Pop")) 
```

#Load the LSOA shape data for EW

```{r}
setwd(LSOAshapedata)

shape <- readOGR(dsn = list.files(pattern = "shp"))

```

# Load Post code data

This chunk checks the exiatane of the postcode to lsoa lookup if it doesn't find it, the second part of the code is run that creates the lookup. Creating the lookup takes quite a lot of time so it is good to only have to do it once.
```{r}

if(file.exists(file.path(DataFolder,"PstCdLSOA.rds"))){
 PstCdLSOA.raw <- readRDS(file.path(DataFolder,"PstCdLSOA.rds"))
} else {
  source(file.path(CommonCode, "PostcodeLookup.R"))
}

```

Bind the ward names and LAD names to the lsoa data
```{r LAD/WARD data}
setwd(PostcodeLookups)
lsoa <- read_csv("PCD11_OA11_LSOA11_MSOA11_LAD11_EW_LU_aligned_v2.csv") %>%
  select(LSOA11CD:LAD11NMW) %>% distinct(., LSOA11CD, .keep_all = TRUE)

setwd(AddGeog)
wardnames <- read_csv("Ward_to_Local_Authority_District_to_County_to_Region_to_Country_December_2016_Lookup_in_United_Kingdom_V2.csv") %>% setNames(c("Admin_ward_code", names(.)[-1]))

#uses the dataframe from the CreateExceltemplates to get the ward names and the lsoa together
wardlsoa <- PstCdLSOA.raw %>% distinct(lsoa11cd, .keep_all = TRUE) %>% 
  select(lsoa11cd, Admin_ward_code)

EW2 <- inner_join(EW, lsoa, by = c("ECODE" = "LSOA11CD") ) %>% 
      left_join(., wardlsoa, by = c("ECODE"="lsoa11cd") ) %>% 
  left_join(., select(wardnames, Admin_ward_code, WD16NM), by = "Admin_ward_code") %>%
  group_by(Admin_ward_code) %>%
  mutate(WardHomes = sum(Homes),  WardPop = sum(Pop)) %>% ungroup
rm(wardlsoa)
rm(wardnames)
```

Convert shape file to dataframe and join in the LSOA data
```{r}

if(file.exists(file.path(DataFolder,"shapeframe.rds"))){
 AG<- readRDS(file.path(DataFolder,"shapeframe.rds"))
} else{
    setwd(LSOAshapedata)

    shape <- readOGR(dsn = list.files(pattern = "shp"))

    AG <- fortify(shape, region = "objectid")

    AG <- AG %>% mutate(id = as.integer(id)) %>% 
      left_join(., shape@data, by = c("id" = "objectid")) %>% 
    left_join(., EW2, by = c("lsoa11cd" = "ECODE"))

    saveRDS(AG, file.path(DataFolder,"shapeframe.rds"))
    }

```

#I should probably join the homes and pop at the same time I join the council tax data

#get prices

```{r}

setwd(DataFolder)

prices <-read_csv("pp-2016.csv", col_names = FALSE ) %>% 
  mutate(X4 = gsub(" ", "", X4)) %>% 
  left_join(., PstCdLSOA.raw %>% mutate(Postcode = gsub(" ", "", Postcode)), 
            by = c("X4"="Postcode")) %>% 
  filter(!is.na(lsoa11cd)) %>% mutate(Prime = X2>1e6, SuperPrime = X2>1e7)

MeanWardPrice <- prices %>% 
  group_by(Admin_ward_code) %>% 
  summarise( MeanPrice = mean(X2), 
             MedianPrice = median(X2),counts =n(),  
             Prime = sum(Prime), 
             SuperPrime = sum(SuperPrime)) %>% 
  mutate(TotPrime = Prime/sum(Prime), TotSupPrime = SuperPrime/sum(SuperPrime),
         PrimePerc = Prime/counts, SuperPrimePerc = SuperPrime/counts)

MeanWardPrice %>% ggplot(., aes(x= counts)) + geom_density()

WardDat2 <- WardDat %>% left_join(., MeanWardPrice) %>% mutate(vallow = LowUse*MeanPrice) %>%
  filter(!is.na(MeanPrice))

setwd(DataFolder)
list.files()

#Unnfordability at EW level salery 2014-2015
mean(prices$X2)/31920

#is mean income estimates
IncomeEst <- read_excel("1smallareaincomeestimatesdataupdate.xls", sheet = 4, skip = 4) %>%
  setNames(make.names(names(.)) %>% 
             gsub("\\.(?=\\.*$)", "", ., perl=TRUE)) #removes trailing full stop.

test2 <- prices %>% left_join(., select(EW2, ECODE, MSOA11CD), 
                              by =c("lsoa11cd"="ECODE")) %>%
  group_by(MSOA11CD) %>%
  summarise(MeanPrice = mean(X2), counts = n()) %>%
  left_join(., IncomeEst, by=c("MSOA11CD"= "MSOA.code")) %>%
  mutate(Yearly.income= 52 * Total.weekly.income, ratio = MeanPrice/Yearly.income)

test2 %>% ggplot(.,aes(x=ratio)) +geom_density()

sum(test2$ratio>6)/nrow(test2)  #about 50% EW
test2 %>% filter(Region.name == "London") %>% mutate(HighRatio = ratio>10) %>%
  summarise(total = sum(HighRatio)/n())

```

# Using the full data

```{r}
#Lambeth

setwd(file.path(RegionsComp, "Region London"))

Lambethinfo <- read_excel("LambethExemptionsLSOA.xlsx")[,1:4] %>%  
    StructureData(., lowuse = c(2:5,24,25), empty =  c(4:5,24,25))

#Islington is exemptions only

Islingtoninfo <- read_csv("IslingtonExemptionsLSOA as at 2017 03 24.csv"  )[1:4]%>%  
    StructureData(., lowuse = c(2:11,14:18))

#Greenwhich

Greenwichinfo <- read_excel( "Royal Borough Of Greenwich.xlsx")[1:4] %>%  
    StructureData(., c(2,3,4,21,22), c(3,4,21,22))

#Camden

Camdeninfo <- read.xlsx("Camden Discounts LSOA - 20891480.xlsx" )[,1:4] %>%
 StructureData()

#Bromley

Bromleyinfo <- read.xlsx("Copy of BromleyExemptionsLSOA.xlsx")[,c(1,4:6)] %>%
  StructureData(c(6:8,29:31))

#Haringey

#Haringey is funney and has to be worked over a bit first
Haringeyinfo <- read.xlsx("Haringey_Exemptions_Discounts_LSOA.xlsx")[c(1,3,5)]

Haringeyinfo <- data_frame(Discount = rep(Haringey$`Exemption./.Discount.type`, Haringey$Count.of.Exemptions.by.Post.Code),
           LSOA = rep(Haringey$LSOA_CODE, Haringey$Count.of.Exemptions.by.Post.Code)) %>%
  mutate( a= NA, b = NA) %>%
  StructureData(21:25)

#Sutton

Suttoninfo <- read_excel("SuttonExemptionsLSOADiscount.xlsx"  )[,c(1,4:6)] %>%
  StructureData()

#Richmond
Richmondinfo <- read_csv("Richmond_2.csv" )[1:4]%>%
  StructureData()

#Hillingdon

Hillingdoninfo <- read.xlsx("HillingdonDiscountsLSOA.XLSX" )[1:4]%>% 
  StructureData()

#Bexley

Bexley <- read.xlsx( "Bexley Exemptions.xlsx" , colNames = FALSE) %>%
StructureData

#RBKC

Chelseainfo <- read_excel( "Kensington and ChelseaDiscountsLSOA.xlsx" )[,1:4] %>%
  StructureData()

#Barking
Barkinginfo <- read.xlsx( "Barking and DagenhamExemptionsLSOA v3.xlsx" , colNames = FALSE)[,1:4] %>%
  StructureData(c(2,3,6,7))

#Barnet

Barnetinfo <- read.xlsx( "3386297 Attachment BarnetDiscounts.xlsx"  , colNames = TRUE) %>%
  StructureData(2:8)

#Newham

Newhaminfo <- read.xlsx( "NewhamExemptionsLSOA.XLSX"  , colNames = TRUE)[,1:4] %>%
  StructureData(2)

test <- ls(pattern = "info") %>% map_df(~eval(parse(text=.x)))

```

#Semi data 
```{r}
#ward folder
setwd(SemiLondon)
list.files()

wardnames <- EW2 %>% group_by(Admin_ward_code, WD16NM) %>% summarise(counts = n(), LAD11NM = first(LAD11NM), LAD11CD = first(LAD11CD)) %>% arrange(WD16NM)

test <- wardnames %>% filter(grepl("Kensington", LAD11NM))

test <- EW2 %>% filter(LAD11NM == "Wandsworth")

#Hackney has provided Census Wards I'm not sure how useable it is. 
hackney <- read_csv("hackney.csv") %>% setNames(make.names(names(.))) %>% group_by(Exemption.Discount, Ward.Name) %>% summarise(counts = n()) %>% ungroup %>% spread(Exemption.Discount, counts, fill = 0)

#Hammersmith 
Hammersmith <- read_excel("Hammersmith.xls", skip = 1 ) %>% filter(!is.na(Ward) ) %>% 
  bind_cols(., filter(wardnames,LAD11NM == "Hammersmith and Fulham")) %>% select(-Ward) %>%
  mutate(LowUse = `Second (B)` + `Empty Unocc`+ LTE + Premium, Empty = LowUse - `Second (B)`)

Hammersmith2 <- EW2 %>% filter(LAD11NM == "Hammersmith and Fulham") %>% left_join(., select(Hammersmith, Admin_ward_code, LowUse, Empty))

HAMDAT <- AG %>% filter(LAD11NM == "Hammersmith and Fulham")%>% left_join(., select(Hammersmith, Admin_ward_code, LowUse, Empty))

#Wandsworth

wardnames %>% filter(LAD11NM == "Wandsworth")

Wandsworth <- read_excel("wandsworth2.xlsx",  skip = 3 ) %>% select(1,3) %>% 
  group_by(`Ptc Wardname`) %>% summarise(Empty = sum(`Count Current Occupancy Discounts`)) %>% ungroup %>% 
  arrange(`Ptc Wardname`) %>% 
  filter(!is.na(`Ptc Wardname`)) %>% 
  bind_cols(., wardnames %>% filter(LAD11NM == "Wandsworth") %>% 
              select(Admin_ward_code)) %>%
  mutate(LowUse = Empty)

Wandsworth2 <- EW2 %>% filter(LAD11NM == "Wandsworth") %>% left_join(., select(Wandsworth, Admin_ward_code, Empty))
WANDAT <- AG %>% filter(LAD11NM == "Wandsworth")%>% left_join(., select(Wandsworth, Admin_ward_code, LowUse, Empty))

#Merton
wardnames %>% filter(LAD11NM == "Merton")
Merton <- read_excel("merton1.xlsx" )[,-2] %>%setNames(make.names(names(.))) %>% slice(-1) %>% arrange(X__1) %>%
  filter(X__1 != "Norbury", X__1!="Nightingale", X__1 !="Stonecot") %>% 
    bind_cols(., wardnames %>% filter(LAD11NM == "Merton") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty =   Disc.Class.C.Vacant.dwelling + Disc.class.C.Long.term.empty + Premium.Levy, LowUse =  Disc.Class.A.Second.home + Disc.Class.B.second.home )

MERDAT <- AG %>% filter(LAD11NM == "Merton")%>% left_join(., select(Merton, Admin_ward_code, LowUse, Empty))

#Southwark

test <- wardnames %>% filter(LAD11NM == "Southwark")

Southwark <- read_excel("southwark.xlsx" , skip = 1) %>% setNames(make.names(names(.))) %>% 
  filter(Ward.codes != "TOTALS") %>% select(-Ward.codes) %>%
  mutate(WARDS= na.locf(WARDS)) %>% group_by(WARDS) %>%
  summarise_all(sum, na.rm = TRUE) %>% ungroup %>%
  arrange(WARDS) %>%
    bind_cols(., wardnames %>% filter(LAD11NM == "Southwark") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty =   Unoccupied.and.furnished.lets....0. + Job.related.dwelling...50..discount + Second.Home...unoccupied.and.furnished...0.+Empty.and.unfurnished...0. + Levy.on.long.term.empty...50..Premium + Uninhabitable.Works.required....0., LowUse =  Empty )

Wandsworth2 <- EW2 %>% filter(LAD11NM == "Southwark") %>% left_join(., select(Southwark, Admin_ward_code, Empty))

SOUDAT <- AG %>% filter(LAD11NM == "Southwark")%>% left_join(., select(Southwark, Admin_ward_code, LowUse, Empty))

#Harrow

test <- read_excel("HarrowCTBandsLSOA. 3118795.xlsx"  )[1:4]%>% setNames(make.names(names(.)))

#Kingston
test <- wardnames %>% filter(grepl("Kingston upon Thames", LAD11NM))

Kingston <- read_excel("Kingston1.xlsx"  )%>% setNames(make.names(names(.))) %>%
  arrange(X__1) %>% slice(1:16) %>%
  bind_cols(., wardnames %>% filter(LAD11NM == "Kingston upon Thames") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty = Empty...Furnished..0.. + Empty.Homes.Premium + LONG.TERM.EMPTY..0.. + Structural.Repair..0.., LowUse = Empty + X2nd.Home..0..)

KINDAT <- AG %>% filter(LAD11NM == "Kingston upon Thames")%>% left_join(., select(Kingston, Admin_ward_code, LowUse, Empty))

#TowerHamlets
test <- wardnames %>% filter(grepl("Tower", LAD11NM))

Tower <- read_excel("Towerhamlets.xlsx")[,c(1,3)] %>% 
  filter(!is.na(DiscClass)) %>%
  group_by(Ward, DiscClass) %>%
  summarise(count = n()) %>% ungroup %>%
  spread(., key =DiscClass, value = count, fill = 0 ) %>% 
  mutate(Empty = rowSums(.[,4:8]), LowUse = B+BE+Empty) %>% 
  slice(c(1:13, 15:17, 14, 18:20)) %>%
   bind_cols(., wardnames %>% filter(grepl("Tower", LAD11NM)))

TOWDAT <- AG %>% filter(grepl("Tower", LAD11NM)) %>% left_join(., select(Tower, Admin_ward_code, LowUse, Empty))

#Havering Is census wards not electoral ones
# test <- wardnames %>% filter(grepl("Havering", LAD11NM))
# 
# Havering <- read.xlsx("Havering1.xlsx")[,-2] %>% gather(key = "Ward", value = "counts",-X1) %>%
#   mutate(X1 = make.names(X1)) %>%
#   spread(key = "X1", value = "counts", fill = 0)

```

Plot Map
```{r}
AG2 <- bind_rows(CHELDAT,HAMDAT, WANDAT, LAMDAT, MERDAT, 
                 SOUDAT, KINDAT, TOWDAT, GREDAT, CAMDAT, 
                 BRODAT, HARDAT, RICDAT, HILDAT, BEXDAT,
                 SUTDAT, ISLDAT, BARKDAT, BARNDAT) %>% 
  mutate(LowUsePerc = LowUse/WardHomes*100, 
         EmptyPerc = Empty/WardHomes*100,
         EmptyPercRank = percent_rank(Empty/WardHomes))

ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group, fill = EmptyPerc), 
colour = alpha("darkred", 1/2), size = 0.1) + scale_colour_gradient2() +
  labs(title = "Current London Data") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank())

SaveFig("London.png")

AG2 %>% 
  ggplot(.)+ geom_polygon( aes(long, lat, group = group, fill =sqrt(EmptyPerc)), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2()

AG2 %>% 
  ggplot(.)+ geom_polygon( aes(long, lat, group = group, fill =EmptyPercRank), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2() +
  labs(title = "Current London Data Coloured by Percent Rank") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank())

SaveFig("LondonPerc.png")

ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group, fill = LowUse/WardHomes), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2()

#single borough

PlotMap(Chelsea, "Chelsea")

SaveFig("RBKCLowUse.png")

ggplotly() #STFD!

```

#Proper data Barrow

```{r}
setwd(file.path(RegionsComp, "Region North West"))

list.files()

Barrow <- read_excel("Barrow-in-FurnessExemptionsLSOA.xlsx", sheet=1) %>% setNames(make.names(names(.))) %>% .[,1:4] %>% filter(Current.Discount.Type.Code !="SINGLE") %>% group_by(LSOA_CODE, Current.Discount.Type.Code) %>%
  summarise(Empty = n()) %>% 
  spread(key=Current.Discount.Type.Code, value = Empty, fill = 0 ) %>% ungroup %>%
 mutate( LowUse = rowSums(.[-1]), Empty = LowUse-PCLB0)

Barrow2 <- Barrow %>% 
  left_join(., EW2, by = c("LSOA_CODE" = "ECODE")) %>% 
  mutate(LowUsePerc = LowUse/Homes, EmptyPerc = Empty/Homes)

Barrow3 <- Barrow %>% left_join(., AG, by = c("LSOA_CODE" = "lsoa11cd")) %>% 
  mutate(LowUsePerc = LowUse/Homes, EmptyPerc = Empty/Homes)

ggplot()+ geom_polygon(data=Barrow2, aes(long, lat, group = group, fill = PCLB0), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2() + ggtitle("Barrow in Furness Emptyness %")
```

```{r}
SLakeland <- read_excel("South LakelandExemptionsLSOA.XLSX", sheet=1)%>% setNames(make.names(names(.))) %>% .[,1:4] %>%
  group_by(Exemption.type, LSOA_CODE) %>% summarise(counts = n()) %>% 
  spread(key= Exemption.type, value = counts, fill = 0) %>% 
  mutate(Empty = rowSums(.[,2:9])) %>%
  left_join(., EW2, by = c("LSOA_CODE" = "ECODE")) %>% mutate(EmptyPerc = Empty/Homes)

SLakeland2 <- SLakeland %>% left_join(., AG, by = c("LSOA_CODE" = "lsoa11cd"))

ggplot()+ geom_polygon(data=SLakeland2, aes(long, lat, group = group, fill = EmptyPerc), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2() + ggtitle("Barrow in Furness Emptyness %")

bind_rows(Barrow3,SLakeland2) %>%ggplot(.)+ geom_polygon( aes(long, lat, group = group, fill = EmptyPerc), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2() + 
  ggtitle("Cumbria Empty Percentage")+
  theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank())

```

#Cornwall

```{r}

setwd(file.path(RegionsComp, "Region South West"))

list.files()

Cornwall <- read_excel("Cornwall.xlsx", sheet=1)%>% setNames(make.names(names(.))) %>% .[,1:4] %>%
  group_by(Exemption.type, LSOA_CODE) %>% summarise(counts = n()) %>% 
  spread(key= Exemption.type, value = counts, fill = 0) %>% 
  mutate(LowUse = rowSums(.[,c(FALSE,!grepl("exempt",(names(.)[-1])))])) %>%
  left_join(., EW2, by = c("LSOA_CODE" = "ECODE")) %>% group_by(Admin_ward_code)  %>%
  mutate(LowUsePerc = LowUse/Homes, WardLowUse = sum(LowUse) ,WardLowUsePerc = WardLowUse/WardHomes) %>% ungroup

Cornwall <- Cornwall %>% left_join(., MeanWardPrice) %>% mutate(ValLow = LowUse*MeanPrice)

C2 <- Cornwall %>% 
  mutate(LowUsePerc = LowUsePerc*100,
         WardLowUsePerc = WardLowUsePerc*100) %>% 
  left_join(., AG, by = c("LSOA_CODE" = "lsoa11cd"))

ggplot(C2, aes(long, lat, group = group, fill = WardLowUsePerc))+ 
  geom_polygon(colour = alpha("black", 1/4), size = 0.2) + 
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = mean(C2$WardLowUsePerc, na.rm = TRUE))+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank())+ 
  labs(title = "Low Use Homes in the County of Cornwall", 
       fill="Percentage")     
SaveFig("Cornwall.png")

CornCor <- Cornwall %>% 
  #filter(Admin_ward_code !="E05009195") %>% 
  summarise(x =cor(MedianPrice, WardLowUsePerc, use = 'complete.obs'))

 Cornwall %>% ggplot(., aes(y=MedianPrice, x= WardLowUsePerc )) + geom_point() +
   labs(title = "Mean Price vs Low Use % by ward in Cornwall",
        x = "Low Use Percentage by Ward", 
        y = "Mean Price") +
  theme(plot.title = element_text(hjust = 0.5)) + 
    annotate("text", 
             label = paste0("Correlation =", signif(CornCor[1,1],2)), 
             x = 0.35 , y = 2.5e5, size = 5) +
   geom_smooth(method = "lm", se = FALSE)#+
#   coord_cartesian(xlim=c(0,0.2), ylim= c(1e5, 3.5e+5))
SaveFig("CornCor.png")

```

#Suffolk

```{r}

setwd(file.path(RegionsComp, "Region East of England"))
list.files()

CoastWave <- bind_rows(read.xlsx("FOI IMT178324 SuffolkCoastal.xlsx")[1:4] , read.xlsx("FOI IMT178324 Waveney.xlsx")[1:4]) %>% setNames(make.names(names(.))) %>%
  mutate(Exemption.type = Exemption.type %>% trimws %>%
           make.names %>% 
           gsub("\\.{2,}" , ".", .)) %>%
  group_by(Exemption.type, LSOA_CODE) %>% summarise(counts = n()) %>% 
  spread(key= Exemption.type, value = counts, fill = 0) %>% 
    setNames(make.names(trimws(names(.)))) %>%
    mutate(LowUse = rowSums(.[,c(5,6,13,14,16:18)]), Empty = LowUse -Second.Home.Class.A -Second.Home.Class.B) %>%
  left_join(., EW2, by = c("LSOA_CODE" = "ECODE")) %>% group_by(Admin_ward_code)  %>%
  mutate(LowUsePerc = LowUse/Homes, WardLowUse = sum(LowUse) ,WardLowUsePerc = WardLowUse/WardHomes) %>% ungroup %>% left_join(., MeanWardPrice) %>% mutate(ValLow = LowUse*MeanPrice)

CoastWave %>% left_join(., AG, by = c("LSOA_CODE" = "lsoa11cd"))  %>% 
  ggplot(.)+ 
  geom_polygon( aes(long, lat, group = group, fill = LowUsePerc), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2()

#corellation
CoastWave %>% 
  #filter(Admin_ward_code !="E05009195") %>% 
  summarise(x =cor(MeanPrice, WardLowUsePerc, use = 'complete.obs'))

 CoastWave %>% ggplot(., aes(y=MeanPrice, x= WardLowUsePerc )) + geom_point() +
   labs(title = "Coastal and Waveny: Relationship between LowUse\n Percentage and Mean Price in Ward")+
  theme(plot.title = element_text(hjust = 0.5))

```

#Hull
```{r}

test <-read_excel(file.path(RegionsComp, "Region Yorkshire and The Humber",
                "KINGSTON UPON HULL%2c CITY OFDISCOUNTSLSOA.XLSX"))[-c(c(1:6)),1:4] %>%
  CleanData(.)

PlotMap(test, AG, variable = "WardLowUsePerc")

```