### 2017-05-29 PlotwithLeaflet.Rmd
Leaflet
Attempting to plto with leaflet to make the maps more intereesting. It is proving very difficult.

---
title: "Untitled"
author: "Jonathan Bourne"
date: "29 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}

setwd("/home/jonno")
list.files()

states <- geojsonio::geojson_read("us-states.geojson" , what = "sp")
class(states)

m <- leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))

```

```{r}
MyData <- AG %>% filter(grepl("kensington", LAD11NM ,ignore.case = T)) %>%
  mutate(group = lsoa11cd)

MyData <- SpatialPointsDataFrame(coords =MyData[,c(1,2)], data = MyData,
                               proj4string = CRS("+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs +ellps=airy +towgs84=446.448,-125.157,542.060,0.1502,0.2470,0.8421,-20.4894"))

#random thing
stuff <-data.frame(lsoa11cd=unique(MyData$lsoa11cd),row.names=unique(MyData$id))

MyData<- points2polygons(MyData,stuff)

plot(spDF,col=spDF$box_id+1)

class(MyData) # data.frame

m <- leaflet(MyData) %>%
  #setView(-96, 37.8, 4) %>%
  addTiles()

  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))

m %>% addPolygons()

```

```{r}
# this code just creates a sample SpatialPointsDataFrame 
x <- c(-10,-10,10,10,-10)
y <- c(-10,10,10,-10,-10)
df.1 <- data.frame(x,y,id=47, order=1:5,hole=F,piece=1,group=47.1,box_id=1)
coordinates(df.1)=c("x","y")
x <- c(+15+3*cos(2*pi/5*(0:5)))
y <- c(-15+3*sin(2*pi/5*(0:5)))
df.2 <- data.frame(x,y,id=48, order=1:6,hole=F,piece=1,group=48.1,box_id=2)
coordinates(df.2)=c("x","y")
x <- c(-15+2*cos(2*pi/8*(0:8)))
y <- c(+15+2*sin(2*pi/8*(0:8)))
df.3.1 <- data.frame(x,y,id=20, order=1:9,hole=F,piece=1,group=20.1,box_id=3)
coordinates(df.3.1)=c("x","y")
x <- c(0+2*cos(2*pi/8*(0:8)))
y <- c(+15+2*sin(2*pi/8*(0:8)))
df.3.2 <- data.frame(x,y,id=20, order=1:9,hole=F,piece=1,group=20.2,box_id=3)
coordinates(df.3.2)=c("x","y")
x <- c(+15+2*cos(2*pi/8*(0:8)))
y <- c(+15+2*sin(2*pi/8*(0:8)))
df.3.3 <- data.frame(x,y,id=20, order=1:9,hole=F,piece=1,group=20.3,box_id=3)
coordinates(df.3.3)=c("x","y")
df <- rbind(df.1,df.2,df.3.1,df.3.2,df.3.3)
df
#              coordinates id order  hole piece group box_id
# 1             (-10, -10) 47     1 FALSE     1  47.1      1
# 2              (-10, 10) 47     2 FALSE     1  47.1      1
# 3               (10, 10) 47     3 FALSE     1  47.1      1
# 4              (10, -10) 47     4 FALSE     1  47.1      1
# 5             (-10, -10) 47     5 FALSE     1  47.1      1
# 6              (18, -15) 48     1 FALSE     1  48.1      2
# 7  (15.92705, -12.14683) 48     2 FALSE     1  48.1      2
# 8  (12.57295, -13.23664) 48     3 FALSE     1  48.1      2
# ...
data <- data.frame(box_id=unique(df$box_id),row.names=unique(df$id))
```

```{r}
points2polygons <- function(df,data) {
  get.grpPoly <- function(group,ID,df) {
         Polygon(coordinates(df[df$id==ID & df$group==group,]))
         }
  get.spPoly  <- function(ID,df) {
         Polygons(lapply(unique(df[df$id==ID,]$group),get.grpPoly,ID,df),ID)
         }
  spPolygons  <- SpatialPolygons(lapply(unique(df$id),get.spPoly,df))
  SpatialPolygonsDataFrame(spPolygons,match.ID=T,data=data)
}

spDF <- points2polygons(df,data)
plot(spDF,col=spDF$box_id+1)

library(rgdal)
writeOGR(spDF,dsn=".",layer="myShapefile", driver="ESRI Shapefile")
```