### 2017-05-29 PlotwithLeaflet.Rmd
LEaflet update
Have managed to make a leaflet doc and save it as an html widget

---
title: "Untitled"
author: "Jonathan Bourne"
date: "29 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}

setwd("/home/jonno")
list.files()

states <- geojsonio::geojson_read("us-states.geojson" , what = "sp")
class(states)

points2polygons <- function(df,data) {
  get.grpPoly <- function(group,ID,df) {
         Polygon(coordinates(df[df$id==ID & df$group==group,]))
         }
  get.spPoly  <- function(ID,df) {
         Polygons(lapply(unique(df[df$id==ID,]$group),get.grpPoly,ID,df),ID)
         }
  spPolygons  <- SpatialPolygons(lapply(unique(df$id),get.spPoly,df))
  SpatialPolygonsDataFrame(spPolygons,match.ID=T,data=data)
}

m <- leaflet(states) %>%
  #setView(-96, 37.8, 4) %>%
  addTiles()
  # addProviderTiles("MapBox", options = providerTileOptions(
  #   id = "mapbox.light",
  #   accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))

m %>% addPolygons()

```

```{r}
MyData <- AG %>% filter(grepl("kensington", LAD11NM ,ignore.case = T)) %>%
  rename(LSOA_CODE = lsoa11cd)

MyData <- SpatialPointsDataFrame(coords =MyData[,c(1,2)], data = MyData,
                               proj4string = CRS('+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs'))
MyData <- spTransform(MyData, CRS("+init=epsg:4326"))
#random thing
stuff <-data.frame(LSOA_CODE=unique(MyData$LSOA_CODE),row.names=unique(MyData$id))

MyData<- points2polygons(MyData,stuff)

AG2 <- AG %>% group_by(LSOA_CODE) %>% summarise_all(funs(first))

MyData <- merge(MyData, Chelseainfo)

class(MyData) # data.frame

v<-"pk.eyJ1Ijoiam9ubm9iIiwiYSI6ImNqM2FlNXhpbDAwM3gzMW83ZW1hbHI0bzYifQ.4AfaY0Un0bKU41hZscpy9Q"

m <- leaflet(MyData) %>%
    #addTiles()
    addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = v))

bins <- c(0, 10, 20, 25, 100)
pal <- colorBin("YlOrRd", domain = MyData$LowUsePerc, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>Low-use  %g &#37",
  MyData$WD16NM, MyData$LowUsePerc
) %>% lapply(htmltools::HTML)

z <-m %>% addPolygons(
  fillColor = ~pal(LowUsePerc),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlight = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
  label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto")) %>%
  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
    position = "bottomright")

saveWidget(z, file="m.html" )

mapshot(z, url = paste0(getwd(), "/map.html"))
```