### 2017-05-30 PlotwithLeaflet.Rmd
Simplify polygon
Used the rmapshaper package to simplify the polygons which means they take up less space than before, still the code can only be used if the other data is loaded in seperate files.

---
title: "Untitled"
author: "Jonathan Bourne"
date: "29 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}

setwd("/home/jonno")
list.files()

states <- geojsonio::geojson_read("us-states.geojson" , what = "sp")
class(states)

points2polygons <- function(df,data) {
  get.grpPoly <- function(group,ID,df) {
         Polygon(coordinates(df[df$id==ID & df$group==group,]))
         }
  get.spPoly  <- function(ID,df) {
         Polygons(lapply(unique(df[df$id==ID,]$group),get.grpPoly,ID,df),ID)
         }
  spPolygons  <- SpatialPolygons(lapply(unique(df$id),get.spPoly,df))
  SpatialPolygonsDataFrame(spPolygons,match.ID=T,data=data)
}

m <- leaflet(states) %>%
  #setView(-96, 37.8, 4) %>%
  addTiles()
  # addProviderTiles("MapBox", options = providerTileOptions(
  #   id = "mapbox.light",
  #   accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))

m %>% addPolygons()

```

```{r}
MyData <- AG %>% filter(grepl("kensington", LAD11NM ,ignore.case = T)) %>%
  rename(LSOA_CODE = lsoa11cd)

MyData <- SpatialPointsDataFrame(coords =MyData[,c(1,2)], data = MyData,
                               proj4string = CRS('+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs'))
MyData <- spTransform(MyData, CRS("+init=epsg:4326"))
#random thing
stuff <-data.frame(LSOA_CODE=unique(MyData$LSOA_CODE),row.names=unique(MyData$id))

MyData<- points2polygons(MyData,stuff)

#AG2 <- AG %>% group_by(LSOA_CODE) %>% summarise_all(funs(first))

MyData <- Chelseainfo %>%
  select(LSOA_CODE, LowUsePerc, WD16NM)%>%
  merge(MyData, .)

class(MyData) # data.frame

v<-"pk.eyJ1Ijoiam9ubm9iIiwiYSI6ImNqM2FlNXhpbDAwM3gzMW83ZW1hbHI0bzYifQ.4AfaY0Un0bKU41hZscpy9Q"

m <- leaflet(MyData) %>%
    #addTiles()
    addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = v))

bins <- c(0, 10, 20, 25, 100)
pal <- colorBin("YlOrRd", domain = MyData$LowUsePerc, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>Low-use  %g &#37",
  MyData$WD16NM, MyData$LowUsePerc
) %>% lapply(htmltools::HTML)

z <-m %>% addPolygons(
  fillColor = ~pal(LowUsePerc),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlight = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
  label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto")) %>%
  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
    position = "bottomright")

#saveWidget(z, file="m.html", selfcontained = FALSE )
setwd("/home/jonno")
mapshot(z, url = paste0(getwd(), "/map.html"))
```

gSimplify
```{r}
test <- gSimplify(MyData,tol= .0001, topologyPreserve = T) %>% 
  SpatialPolygonsDataFrame(.)

library(geojsonio)
states_json <- geojson_json(MyData,geometry = "polygon")

test <- states_json %>% 
  ms_simplify(keep = 0.005,keep_shapes = TRUE, explode = TRUE) %>% # Simplify polygon
  geojson_sp()

plot(test)

m2 <- leaflet(test) %>%
    #addTiles()
    addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = v))

bins <- c(0, 10, 20, 25, 100)
pal <- colorBin("YlOrRd", domain = test$LowUsePerc, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>Low-use  %g &#37",
  test$WD16NM, test$LowUsePerc
) %>% lapply(htmltools::HTML)

z2 <-m2 %>% addPolygons(
  fillColor = ~pal(LowUsePerc),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlight = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
  label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto")) %>%
  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
    position = "bottomright")
saveWidget(z2, file="mapsimple2.html", selfcontained = FALSE )
mapshot(z2, url = paste0(getwd(), "/mapsimple.html"))
```