### 2017-07-06 LowUseLondon.Rmd
All kinds of debug

---
title: "Empty Homes V1"
author: "Jonathan Bourne"
date: "10 February 2017"
output: html_document
---

man-whitney

It would have been sensible to separate council and private housing

PCLB0 is second homes, PCLC0 is long term empty, PCLC100 is empty and unfurnished up to six months and PCLD100 is undergoing major repairs up to 12 months.

#Telegraph article on pensioners requireing mortgages
http://www.telegraph.co.uk/personal-banking/mortgages/pensioner-mortgages-boom-britains-broken-housing-market/

Mean household income 31920
https://www.ons.gov.uk/peoplepopulationandcommunity/personalandhouseholdfinances/incomeandwealth/datasets/nowcastinghouseholdincomeintheuk

#savils definitions of prime also there segmentation of the market
http://pdf.euro.savills.co.uk/spotlight-on/spotlight-on-london-housing-supply.pdf

```{r setup}
setwd( "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode")
source("Setup.R")

setwd(Functions)
list.files() %>% map(source)
```

```{r}
source(file.path(CommonCode, "AuxdataLoad.R"))
```

# Using the full data

```{r}
#Source the code from another file as it is very long
source(file.path(CommonCode, "LondonFullProcesss.R"))

#All London With missing Boroughs
PlotMap(Hammersmithinfo, AG)

 test <- list("Hammersmithinfo", "Chelseainfo") %>% map_df(~eval(parse(text=.x)))

 Hammersmithinfo %>% filter(LSOA_CODE!="E01001866") %>% 
   bind_rows(Chelseainfo) %>%
   MakeLeaflet(.)

 PlotMap(test, AG)
test %>% MakeLeaflet(.)

test2 <- filter(test, LSOA_CODE=="E01001866")

```

#Chelsea Bootstrapped ratios
```{r}

#percent of chelsea low-use
sum(Chelseainfo$LowUse)/sum(Chelseainfo$Homes)

#low-use value chelsea
sum(v$MeanPrice*v$LowUse)
ChelMeanPrice <- (sum(v$MeanPrice*v$counts)/sum(v$counts))

chelstrap <-  Chelseainfo %>% filter(LAD11CD=="E09000020") %>% 
  DistribCompareBootstrapper(., 1652, 1000)

chelstrap %>% mutate(RatioExvsAct = (RatioExvsAct-1)*100)%>% 
  ggplot(., aes(x= class, y = RatioExvsAct, fill = class)) + geom_boxplot() + 
  labs(title = "Bootstrapped difference from expected number of low-use homes", y = "% difference from expected")
SaveFig("DistributionPlot.png")

 Chelttest <- ValueTtest(chelstrap)

 #boxplot comparison
Chelttest$data.frame %>%
  select(HomesPrice, LowUsePrice) %>%
  gather(key ="type", value = "MeanPrice") %>%
ggplot(., aes(x= type, y = MeanPrice, fill = type)) + geom_boxplot()

#density plot comparison
Chelttest$data.frame %>%
  select(HomesPrice, LowUsePrice) %>%
  gather(key ="type", value = "MeanPrice") %>%
ggplot(., aes(x = MeanPrice, fill = type)) + geom_density()

```

#Semi data

The LSOA will be assigned empty and low use homes by percentage

```{r}
#ward folder
setwd(SemiLondon)
list.files()

wardnames <- EW2 %>% group_by(Admin_ward_code, WD16NM) %>% summarise(counts = n(), LAD11NM = first(LAD11NM), LAD11CD = first(LAD11CD)) %>% arrange(WD16NM)

test <- wardnames %>% filter(grepl("Kensington", LAD11NM))

test <- EW2 %>% filter(LAD11NM == "Wandsworth")

#Hackney has provided Census Wards I'm not sure how useable it is. 
hackney <- read_csv("hackney.csv") %>% setNames(make.names(names(.))) %>% group_by(Exemption.Discount, Ward.Name) %>% summarise(counts = n()) %>% ungroup %>% spread(Exemption.Discount, counts, fill = 0)

# #Hammersmith 
# Hammersmithinfo <- read_excel("Hammersmith.xls", skip = 1 ) %>% filter(!is.na(Ward) ) %>% 
#   bind_cols(., filter(wardnames,LAD11NM == "Hammersmith and Fulham")) %>% select(-Ward) %>%
#   mutate(LowUse = `Second (B)` + `Empty Unocc`+ LTE + Premium, Empty = LowUse - `Second (B)`) %>%
#   SemiDataStructure

#Wandsworth

Wandsworthinfo <- read_excel("wandsworth2.xlsx",  skip = 3 ) %>% select(1,3) %>% 
  group_by(`Ptc Wardname`) %>% summarise(Empty = sum(`Count Current Occupancy Discounts`)) %>% ungroup %>% 
  arrange(`Ptc Wardname`) %>% 
  filter(!is.na(`Ptc Wardname`)) %>% 
  bind_cols(., wardnames %>% filter(LAD11NM == "Wandsworth") %>% 
              select(Admin_ward_code)) %>%
  mutate(LowUse = Empty) %>%
  SemiDataStructure

#Merton

Mertoninfo <- read_excel("merton1.xlsx" )[,-2] %>%setNames(make.names(names(.))) %>% slice(-1) %>% arrange(X__1) %>%
  filter(X__1 != "Norbury", X__1!="Nightingale", X__1 !="Stonecot") %>% 
    bind_cols(., wardnames %>% filter(LAD11NM == "Merton") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty =   Disc.Class.C.Vacant.dwelling + Disc.class.C.Long.term.empty + Premium.Levy, LowUse =  Disc.Class.A.Second.home + Disc.Class.B.second.home ) %>%
  SemiDataStructure

#Harrow

test <- read_excel("HarrowCTBandsLSOA. 3118795.xlsx"  )[1:4]%>% setNames(make.names(names(.)))

#Kingston

Kingstoninfo <- read_excel("Kingston1.xlsx"  )%>% setNames(make.names(names(.))) %>%
  arrange(X__1) %>% slice(1:16) %>%
  bind_cols(., wardnames %>% filter(LAD11NM == "Kingston upon Thames") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty = Empty...Furnished..0.. + Empty.Homes.Premium + LONG.TERM.EMPTY..0.. + Structural.Repair..0.., LowUse = Empty + X2nd.Home..0..) %>%
  SemiDataStructure

#TowerHamlets
Towerinfo <- read_excel("Towerhamlets.xlsx")[,c(1,3)] %>% 
  filter(!is.na(DiscClass)) %>%
  group_by(Ward, DiscClass) %>%
  summarise(count = n()) %>% ungroup %>%
  spread(., key =DiscClass, value = count, fill = 0 ) %>% 
  mutate(Empty = rowSums(.[,4:8]), LowUse = B+BE+Empty) %>% 
  slice(c(1:13, 15:17, 14, 18:20)) %>%
   bind_cols(., wardnames %>% filter(grepl("Tower", LAD11NM))) %>%
  SemiDataStructure

#Havering Is census wards not electoral ones
# test <- wardnames %>% filter(grepl("Havering", LAD11NM))
# 
# Havering <- read.xlsx("Havering1.xlsx")[,-2] %>% gather(key = "Ward", value = "counts",-X1) %>%
#   mutate(X1 = make.names(X1)) %>%
#   spread(key = "X1", value = "counts", fill = 0)

```

Value of low*use and total low use
```{r}

test <-  ls(pattern = "info") %>% map_df(~eval(parse(text=.x))) %>%
  group_by(Admin_ward_code) %>%
  summarise(LAD11NM = first(LAD11NM),
            MeanPrice = first(MeanPrice),
            WardLowUse = first(WardLowUse)) %>%
  group_by(LAD11NM) %>%
  summarise(totalLow = sum(MeanPrice*WardLowUse)/1e6,
            TotalEmpt = sum(WardLowUse))

sum(test$totalLow, na.rm = T)  
sum(test$TotalEmpt, na.rm = T)  
length(unique(test$LAD11NM))

```

Plot Map
```{r}

 ls(pattern = "info") %>% map_df(~eval(parse(text=.x))) %>%
  PlotMap(., filter(AG, RGN15NM=="London"), 
          filtermap = FALSE,
          title = "Map of London with all current data")
SaveFig("LondonCont.png")

 ls(pattern = "info") %>% map_df(~eval(parse(text=.x))) %>%
  PlotMap(., filter(AG, RGN15NM=="London"), variable = "LowuseClass", 
          filtermap = FALSE, 
          title = "Map of London by low-use class")
SaveFig("LondonDisc.png")

#single borough

Chelseainfo %>% mutate(Value = ValLow/1e6) %>%
    filter(LAD11NM == "Kensington and Chelsea") %>%
  PlotMap(., AG, variable = "Value", title = "Low-use homes in Kensington and Chelsea")

Chelseainfo %>% 
  filter(LAD11NM == "Kensington and Chelsea") %>%
  group_by(Admin_ward_code) %>%
  mutate(Value = sum(ValLow)/1e6) %>%
  PlotMap(., AG, variable = "Value", title = "Low-use homes in Kensington and Chelsea")
SaveFig("RBKCvalue.png")

Chelseainfo %>% mutate(Value = ValLow/1e6) %>%
    filter(LAD11NM == "Kensington and Chelsea") %>%
PlotMap(., AG, title = "Low-use homes in Kensington and Chelsea")
SaveFig("RBKCLowUse.png")

PlotMap(CoLinfo, AG, title = "Low-use homes in the city of London")
SaveFig("CoLLowUse.png")

PlotMap(Lambethinfo, AG, title = "Low-use homes in Lambeth")
SaveFig("LamLowUse.png")

ggplotly() #STFD!

corplotter(Chelseainfo, "MeanPrice", "WardLowUsePerc", "Mean Price vs Low Use % by ward in RBKC",
           "Low Use Percentage by Ward",
           "Mean Price")
SaveFig("ChelseaCor.png")

Chelseainfo %>% filter(WardLowUsePerc != max("WardLowUsePerc"),
                       MeanPrice != max(MeanPrice)) %>%
  corplotter(., "MeanPrice", "WardLowUsePerc", "Median Price vs Low Use % by ward in RBKC",
           "Low Use Percentage by Ward",
           "Median Price")
```

```{r}

LambethStrap<-  Lambethinfo %>% filter(LAD11CD == "E09000022") %>% 
  DistribCompareBootstrapper(., 1652, 100)

LambethStrap %>%
mutate(RatioExvsAct = (RatioExvsAct-1)*100)%>% 
  ggplot(., aes(x= class, y = RatioExvsAct, fill = class)) + geom_boxplot() + 
  labs(title = "Bootstrapped difference from expected number of low-use homes", y = "% difference from expected")

#Why are lambteths results junk?
LamStrap <- ValueTtest(LambethStrap)

df <- LambethStrap

CamdenStrap<-  DistribCompareBootstrapper(Camdeninfo, 1652, 1000)

CamdenStrap %>%
mutate(RatioExvsAct = (RatioExvsAct-1)*100)%>% 
  ggplot(., aes(x= class, y = RatioExvsAct, fill = class)) + geom_boxplot() + 
  labs(title = "Bootstrapped difference from expected number of low-use homes", y = "% difference from expected")

#Why are lambteths results junk?
Camttest<- ValueTtest(CamdenStrap)

PlotMap(Camdeninfo,AG)

```