### 2017-07-23 Average local prices.Rmd
Average local prices
Wrote a script to calculate the price of a LAD relative to the local population of LADs... it doesn't improve any models

---
title: "Untitled"
author: "Jonathan Bourne"
date: "23 July 2017"
output: html_document
---

This file find out whether an are is more or less expensive than the local average, where local is defined as the number of x hops from the root node in the LAD network of main land UK

source: http://geoportal.statistics.gov.uk/datasets/osni-open-data-largescale-boundaries-local-government-districts-2012

```{r}
library(shp2graph)

setwd( "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode")
source("Setup.R")

setwd(Functions)
list.files() %>% map(source)

```

```{r}

NILADshapedata <- file.path(basewd,"ShapeFiles","OSNI_Open_Data_Largescale_Boundaries__Local_Government_Districts_2012")
setwd(NILADshapedata)

LADshape <- readOGR(dsn = list.files(pattern = "shp"))

Adjmat <- Shape2Adjmat(LADshape, "LGDCode")

g1 <- graph_from_adjacency_matrix(Adjmat, mode = "undirected")

plot(g1)

BFS1 <- bfs(g1, root = 1, unreachable = FALSE, rank = TRUE, dist = TRUE)

BFS1$dist

NetworkDist <- 1:length(V(g1)) %>% 
  map(~{
  bfs(g1, root = .x, unreachable = FALSE, rank = TRUE, dist = TRUE)$rank %>%
      as.matrix %>% t %>% as.data.frame
  }) %>% 
  bind_rows()

test <- prices %>% filter(grepl("N", Admin_district_code))
  filter(Admin_district_code == "N09000001")

```

```{r}

LADshapedata <- file.path(basewd,"ShapeFiles","Local_Authority_Districts_December_2016_Full_Extent_Boundaries_in_Great_Britain")
setwd(LADshapedata)

LADshape <- readOGR(dsn = list.files(pattern = "shp"))

Adjmat <- Shape2Adjmat(LADshape, "lad16cd")

g1 <- graph_from_adjacency_matrix(Adjmat, mode = "undirected")

#plot the graph
plot(g1)

#Construct a distance network across all the LADs
NetworkDist <- 1:length(V(g1)) %>% 
  map(~{
  bfs(g1, root = .x, unreachable = FALSE, rank = TRUE, dist = TRUE)$rank %>%
      as.matrix %>% t %>% as.data.frame
  }) %>% 
  bind_rows() %>% mutate(LAD11CD = names(.))

LADprice <- prices %>%
  filter(X5!= "O") %>%
  group_by(Admin_district_code) %>%
summarise(price = mean(X2)) %>%
  mutate(meandiff = price-mean(price),
         pricepercent = percent_rank(price))

#Filter the data to only LADs that I have data for

LocalMean <- NetworkDist %>% 
  gather(key= "LAD11CD2", value = "distance", -LAD11CD) %>%
  filter(LAD11CD %in% CombineAB$LAD11CD, distance <=3) %>%
#join on the average prices per LAD
  left_join(LADprice, by = c("LAD11CD2"="Admin_district_code")) %>%
  #find local mean
  group_by(LAD11CD) %>%
  summarise(LocalMean = mean(price,na.rm = TRUE))

test <- Combine1ExO2 %>%
  left_join(LADprice, by = c("LAD11CD"="Admin_district_code"))%>%
  #filter out wales
  filter(!grepl("W", LAD11CD))%>%
  select(LAD11NM, absdiff, LUPdepExt, Extent, maxPerc,Ttest.p.value, pricepercent, LAD11CD) %>%
  mutate(outcome = LUPdepExt>Extent)  %>% 
  left_join(LocalMean, by = "LAD11CD") %>%
  left_join(LADprice, by = c("LAD11CD"="Admin_district_code")) %>%
  mutate(priceRatio = (price/LocalMean)) #%>%
  #select(LAD11CD, LAD11NM, priceRatio, price, LocalMean, absdiff, Ttest.p.value, maxPerc, LUPdepExt)

test %>%
  ggplot(.,aes(x= maxPerc, y = Ttest.p.value, colour = priceRatio>1)) + geom_jitter()

rpart.model <- rpart(as.factor(absdiff>0) ~ maxPerc, test)
rpart.model1 <- rpart(as.factor(absdiff>0) ~ LUPdepExt+ maxPerc, test)
rpart.model2 <- rpart(as.factor(absdiff>0) ~ priceRatio+ maxPerc, test)
rpart.model3 <- rpart(as.factor(absdiff>0) ~ LUPdepExt+ maxPerc + priceRatio, test)

confusionMatrix( predict(rpart.model, test, type = "class"),as.factor(test$absdiff>0))
confusionMatrix( predict(rpart.model1, test, type = "class"),as.factor(test$absdiff>0))
confusionMatrix( predict(rpart.model2, test, type = "class"),as.factor(test$absdiff>0))
confusionMatrix( predict(rpart.model3, test, type = "class"),as.factor(test$absdiff>0))

```