### 2017-08-15 Average local prices.Rmd
New postcodes
New postcodes have been included to make the LSOA matching better.

Also loads more local regions have been added and modelling of the negative positive split.

---
title: "Untitled"
author: "Jonathan Bourne"
date: "23 July 2017"
output: html_document
---

This file find out whether an are is more or less expensive than the local average, where local is defined as the number of x hops from the root node in the LAD network of main land UK

source: http://geoportal.statistics.gov.uk/datasets/osni-open-data-largescale-boundaries-local-government-districts-2012

#this uses rank I think it should be dist check this

```{r}
library(shp2graph)

setwd( "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode")
source("Setup.R")

setwd(Functions)
list.files() %>% map(source)

```

Test on northern Ireland
```{r}

NILADshapedata <- file.path(basewd,"ShapeFiles","OSNI_Open_Data_Largescale_Boundaries__Local_Government_Districts_2012")
setwd(NILADshapedata)

LADshape <- readOGR(dsn = list.files(pattern = "shp"))

Adjmat <- Shape2Adjmat(LADshape, "LGDCode")

g1 <- graph_from_adjacency_matrix(Adjmat, mode = "undirected")

plot(g1)

BFS1 <- bfs(g1, root = 1, unreachable = FALSE, rank = TRUE, dist = TRUE)

BFS1$dist

NetworkDist <- 1:length(V(g1)) %>% 
  map(~{
  bfs(g1, root = .x, unreachable = FALSE, rank = TRUE, dist = TRUE)$rank %>%
      as.matrix %>% t %>% as.data.frame
  }) %>% 
  bind_rows()

test <- prices %>% filter(grepl("N", Admin_district_code))
  filter(Admin_district_code == "N09000001")

```

```{r}

LADshapedata <- file.path(basewd,"ShapeFiles","Local_Authority_Districts_December_2016_Full_Extent_Boundaries_in_Great_Britain")
setwd(LADshapedata)

LADshape <- readOGR(dsn = list.files(pattern = "shp"))

Adjmat <- Shape2Adjmat(LADshape, "lad16cd")

g1 <- graph_from_adjacency_matrix(Adjmat, mode = "undirected")

#plot the graph
plot(g1)

#Construct a distance network across all the LADs
#dist now fixed!
NetworkDist <- 1:length(V(g1)) %>% 
  map(~{
  bfs(g1, root = .x, unreachable = FALSE, rank = TRUE, dist = TRUE)$dist %>%
      as.matrix %>% t %>% as.data.frame
  }) %>% 
  bind_rows() %>% mutate(LAD11CD = names(.))

LADprice <- prices %>%
  filter(X5!= "O") %>%
  group_by(Admin_district_code) %>%
summarise(price = mean(X2)) %>%
  mutate(meandiff = price-mean(price),
         pricepercent = percent_rank(price))

#Filter the data to only LADs that I have data for

LocalMean <- NetworkDist %>% 
  gather(key= "LAD11CD2", value = "distance", -LAD11CD) %>%
  filter(LAD11CD %in% CombineAB$LAD11CD, distance <=3) %>%
#join on the average prices per LAD
  left_join(LADprice, by = c("LAD11CD2"="Admin_district_code")) %>%
  filter(LAD11CD=="E07000029")

#three or less distance
LocalMean <- NetworkDist %>% 
  gather(key= "LAD11CD2", value = "distance", -LAD11CD) %>%
  filter(LAD11CD %in% CombineAB$LAD11CD, distance <=4) %>%
#join on the average prices per LAD
  left_join(LADprice, by = c("LAD11CD2"="Admin_district_code")) %>%
  #find local mean
  group_by(LAD11CD) %>%
  summarise(LocalMean = mean(price,na.rm = TRUE))

test <- Combine1ExO2 %>%
  left_join(LADprice, by = c("LAD11CD"="Admin_district_code"))%>%
  #filter out wales
  filter(!grepl("W", LAD11CD))%>%
  select(LAD11NM, absdiff, LUPdepExt, Extent, maxPerc,Ttest.p.value,  LAD11CD) %>%
  mutate(outcome = LUPdepExt>Extent)  %>% 
  left_join(LocalMean, by = "LAD11CD") %>%
  left_join(LADprice, by = c("LAD11CD"="Admin_district_code")) %>%
  mutate(priceRatio = (price/LocalMean)) %>%
  mutate(LUPdepExt.Extent = LUPdepExt-Extent, price.LocalMean = price-LocalMean ) #%>%
#select(LAD11CD, LAD11NM, priceRatio, price, LocalMean, absdiff, Ttest.p.value, maxPerc, LUPdepExt)
#remove barrow
#filter(LAD11CD !="E07000027")

test %>%
  filter(maxPerc<10) %>%
  ggplot(.,aes(x= maxPerc, y = Ttest.p.value, colour = priceRatio>1)) + geom_jitter()

test %>%
  filter(maxPerc<10) %>%
  ggplot(.,aes(x= Extent, y = LUPdepExt, colour = absdiff>0)) + geom_jitter()

test2 <- test %>% 
  select(LAD11NM, absdiff, LUPdepExt, Extent, LocalMean, price, priceRatio, maxPerc) %>%
  mutate(LUPdepExt.Extent = LUPdepExt>Extent, price.LocalMean = price>LocalMean ) %>%
  filter( price.LocalMean == FALSE)

table(test$maxPerc>5, test$LUPdepExt.Extent>0, test$absdiff>0)

rpart.model <- rpart(as.factor(absdiff>0) ~ maxPerc, test)
rpart.model1 <- rpart(as.factor(absdiff>0) ~ LUPdepExt+ maxPerc, test)
rpart.model2 <- rpart(as.factor(absdiff>0) ~ LUPdepExt + maxPerc + priceRatio, test)
rpart.model3 <- svm(as.factor(absdiff>0) ~ LUPdepExt+Extent + maxPerc + priceRatio, test, type = "C", kernel = "radial")
rpart.model3 <- svm(as.factor(absdiff>0) ~ maxPerc + LUPdepExt+ Extent, test, type = "C", kernel = "radial")

confusionMatrix( predict(rpart.model, test, type = "class"),as.factor(test$absdiff>0))
confusionMatrix( predict(rpart.model1, test, type = "class"),as.factor(test$absdiff>0))
confusionMatrix( predict(rpart.model2, test, type = "class"),as.factor(test$absdiff>0))
confusionMatrix( predict(rpart.model3, test, type = "class"),as.factor(test$absdiff>0))

test2 <- test[predict(rpart.model2, test, type = "class")!=as.factor(test$absdiff>0),]

```