### 2017-08-15 testCode.Rmd

---
title: "Untitled"
author: "Jonathan Bourne"
date: "10 August 2017"
output: html_document
---

```{r}

CalcLowUse<- function(df){ 
  df %>%
  select(LSOA_CODE, LowUse) %>%
        left_join(., EW2, by = c("LSOA_CODE" = "ECODE")) %>% 
        group_by(Admin_ward_code)  %>%
        mutate(LowUsePerc = round(LowUse/Homes *100), 
               WardLowUse = sum(LowUse) ,
               WardLowUsePerc = round(WardLowUse/sum(Homes)*100)) %>% 
        ungroup %>% 
        left_join(., MeanWardPrice, by = "Admin_ward_code") %>% 
        mutate(ValLow = LowUse*MeanPrice,
               LowuseClass = cut(LowUsePerc, c(-1,10,20, 30,100), 
                                 labels =  c("0-10","11-20","21-30", "30+")),
               WardLowuseClass = cut(WardLowUsePerc, c(0,10,20, 30, 100), 
                                     labels = c("0-10","11-20","21-30", "30+")) )
}

test <- ChelseaLONDATA

test <- ls(pattern = "LON") %>% 
  map_df(~get(.x)) %>%
   group_by(LSOA_CODE) %>%
   summarise(LowUse = sum(LowUse, na.rm = TRUE)) %>% 
  CalcLowUse()

z<- MakeLeaflet(test)

test <-list( WestminsterLONDATA, ChelseaLONDATA, CoLLONDATA, CamdenLONDATA) %>%
  map_df(~.x) %>%   
  group_by(LSOA_CODE) %>%
   summarise(LowUse = sum(LowUse, na.rm = TRUE)) %>% 
  CalcLowUse()

z<- MakeLeaflet(test)

z

```

https://voaratinglists.blob.core.windows.net/html/rlidata.htm

```{r}

setwd(file.path(basewd,"VOA data"))

file2load <-"uk-englandwales-ndr-2017-summaryvaluations-compiled-epoch-0002-baseline-csv.csv"

test0<- read_lines(file2load, n_max=10)

test <- read_delim(file2load, delim = "*", col_names = F)

test2 <- test$`Offices And Premises` %>% table %>% as.data.frame

g1 <- test %>% filter(X1=="01")
g2 <- test %>% filter(X1=="02")

test3 <- g1$X5 %>% table %>% as.data.frame
test4 <- g2$X4 %>% table %>% as.data.frame

```

```{r}
  file2load2 <- "uk-englandwales-ndr-2017-listentries-compiled-epoch-0002-baseline-csv.csv"
test0<- read_lines(file2load2, n_max=10)

CorePstCd <- PstCdLSOA.raw %>%
  left_join(EW2, by = c("lsoa11cd" = "ECODE")) %>%
  select(Postcode, lsoa11cd, LAD11CD, LAD11NM, Admin_ward_code, WD16NM)

CompanyData <- read_delim(file2load2, delim = "*", col_names = F) %>%
  mutate(X15 = gsub(" ", "", X15)) %>%
  left_join(., CorePstCd, by = c("X15" = "Postcode"))

LSOABiz <- CompanyData %>%
  filter(!is.na(LAD11CD)) %>%
  mutate(X6 = gsub("( AND PREMISES)|( & PREMISES)", "",X6),
         X6 = gsub("(AND PREMISES)|(& PREMISES)", "",X6)) %>%
  group_by(lsoa11cd, X6) %>%
  summarise(counts = n(),
            LAD11CD = first(LAD11CD),
            LAD11NM = first(LAD11NM),
            Admin_ward_code = first(Admin_ward_code),
            WD16NM = first(WD16NM))

FirmCounts <- LSOABiz %>%
  group_by(X6) %>%
  summarise(counts = n()) %>%
  arrange(-counts)

Dom <- FirmCounts %>%
  filter(grepl("DOMESTIC",X6))

HOLS <- FirmCounts %>%
  filter(grepl("HOLIDAY",X6))

Catering <- FirmCounts %>%
  filter(grepl("SELF CATER",X6))

```

Look at Westminster

```{r}

West <- LSOABiz %>%
  filter(LAD11CD=="E09000033")

Catering <- West %>%
  filter(grepl("SELF CATER",X6))

test <- list(HammersmithLONDATA, WestminsterLONDATA, ChelseaLONDATA, CoLLONDATA, CamdenLONDATA, LambethLONDATA) %>% 
  map_df(~.x) %>%
  group_by(LAD11CD) %>%
  summarise(
    LAD11NM =first(LAD11NM),
    LowUse = sum(LowUse, na.rm= T),
    Homes = sum(Homes, na.rm= T),
    Pop = sum(Pop, na.rm= T)
            ) %>%
  filter(Homes >2000) %>%
  mutate(LowUseH = LowUse/Homes,
         LowUseP = LowUse/Pop)

```