### 2017-09-05 allEnglandMap.Rmd
Cascading
Addede code to being trying to cascade. Included two new functions to smooth the code. the first identifies validnodes for the injection vector. the second balances the generation and demand of the valid nodes across islolated graph componants

---
title: "Untitled"
author: "Jonathan Bourne"
date: "20 August 2017"
output: html_document
---

```{r}

if(file.exists(file.path(DataFolder,"LADShape.rds"))){
LADShape<- readRDS(file.path(DataFolder,"LADShape.rds"))
} else{
setwd(file.path(basewd,"ShapeFiles", "Local_Authority_Districts_December_2014_Ultra_Generalised_Clipped_Boundaries_in_Great_Britain"))

ladshape <- readOGR(dsn = list.files(pattern = "shp"))

LADShape <- fortify(ladshape, region = "objectid")

LADShape <- LADShape %>% mutate(id = as.integer(id)) %>% 
left_join(., ladshape@data, by = c("id" = "objectid")) #%>% 
#left_join(., EW2, by = c("lsoa11cd" = "ECODE"))

saveRDS(LADShape, file.path(DataFolder,"LADShape.rds"))
}

```

```{r}

LADShape <- LADShape %>%
  filter(grepl("E",lad14cd))

LADShape2 <- LADShape %>% left_join(EnglandComp, by = c("lad14cd" = "LAD11CD"))

```

```{r}
  #These two functions are necessary for Points to polygons
  get.grpPoly <- function(group,ID,df) {
    Polygon(coordinates(df[df$id==ID & df$group==group,]))
  }
  get.spPoly  <- function(ID,df) {
    Polygons(lapply(unique(df[df$id==ID,]$group),get.grpPoly,ID,df),ID)
  }

  #this function is necessary for the rest of the function
  points2polygons <- function(df,data) {
    spPolygons  <- SpatialPolygons(lapply(unique(df$id),get.spPoly,df))
    SpatialPolygonsDataFrame(spPolygons,match.ID=T,data=data)
  }

   MyData <- LADShape2 %>% #filter(AG$LAD11CD %in%  unique(dataframe$LAD11CD) ) %>%
    rename(LAD11CD = lad14cd)

  MyData <- SpatialPointsDataFrame(coords =MyData[,c(1,2)], data = MyData,
                                   proj4string = CRS('+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs'))
  MyData <- spTransform(MyData, CRS("+init=epsg:4326"))
  #random thing
  stuff <-data.frame(LAD11CD=unique(MyData$LAD11CD),row.names=unique(MyData$id))

  MyData<- points2polygons(MyData,stuff)

  #AG2 <- AG %>% group_by(LSOA_CODE) %>% summarise_all(funs(first))

  MyData <- EnglandComp %>%
    select(LAD11CD, Significant.Competition, LAD11NM, Comp.Type) %>%
    merge(MyData, .)

  class(MyData) # data.frame

  v<-"pk.eyJ1Ijoiam9ubm9iIiwiYSI6ImNqM2FlNXhpbDAwM3gzMW83ZW1hbHI0bzYifQ.4AfaY0Un0bKU41hZscpy9Q"

  m <- leaflet(MyData) %>%
    #addTiles()
    addProviderTiles("MapBox", options = providerTileOptions(
      id = "mapbox.light",
      accessToken = v))

  pal <- colorFactor("Accent", domain = MyData$Comp.Type)

  labels <- sprintf(
    "<strong>%s</strong><br/>Market affected %s",
    MyData$LAD11NM, MyData$Comp.Type
  ) %>% lapply(htmltools::HTML)

  z <-m %>% addPolygons(
    fillColor = ~pal(Comp.Type),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")
    ) %>%
    addLegend(pal = pal, values = ~Comp.Type, opacity = 0.7, title = NULL,
              position = "bottomright")

```

```{r}
  z

  mapshot(z, file = file.path(Figures, "EnglandMarket.pdf"))
```