### 2017-09-25 testCode.Rmd
VOA and offshore
made loads of regex for address cleaning etc

---
title: "Untitled"
author: "Jonathan Bourne"
date: "10 August 2017"
output: html_document
---

```{r setup}
setwd( "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode")
source("Setup.R")

setwd(Functions)
list.files() %>% map(source)

Figures <- "/home/jonno/Dropbox/Apps/ShareLaTeX/Empty Homes Write up/Figures"
source(file.path(CommonCode, "AuxdataLoad.R"))
```

```{r}

CombineLowUse<- function(df){ 
  df %>%
  select(LSOA_CODE, LowUse) %>%
    CalcLowUse()

}

test <- ChelseaLONDATA

test <- ls(pattern = "CUMBRIA") %>% 
  map_df(~get(.x)) %>%
   group_by(LSOA_CODE) %>%
   summarise(LowUse = sum(LowUse, na.rm = TRUE)) %>% 
  CalcLowUse()

z<- MakeLeaflet(test)

testTower <- TowerLONDATA %>%
  CalcLowUse()

test2 <- ls(pattern = "LON") %>%
  map_df(~get(.x)) %>%   
  group_by(LSOA_CODE) %>%
   summarise(LowUse = sum(LowUse, na.rm = TRUE)) %>% 
  CalcLowUse()

test2 <-list( WestminsterLONDATA, ChelseaLONDATA, CoLLONDATA, CamdenLONDATA, TowerLONDATA) %>%
  map_df(~.x) %>%   
  group_by(LSOA_CODE) %>%
   summarise(LowUse = sum(LowUse, na.rm = TRUE)) %>% 
  CalcLowUse()

y<- MakeLeaflet(test2)

y

y2<- MakeLeaflet(TowerLONDATA)

y2

```

https://voaratinglists.blob.core.windows.net/html/rlidata.htm

```{r}

setwd(file.path(basewd,"VOA data"))

file2load <-"uk-englandwales-ndr-2017-summaryvaluations-compiled-epoch-0002-baseline-csv.csv"

test0<- read_lines(file2load, n_max=10)

test <- read_delim(file2load, delim = "*", col_names = F)

test2 <- test$`Offices And Premises` %>% table %>% as.data.frame

g1 <- test %>% filter(X1=="01")
g2 <- test %>% filter(X1=="02")

test3 <- g1$X5 %>% table %>% as.data.frame
test4 <- g2$X4 %>% table %>% as.data.frame

```

#VOA stuff

```{r}
  file2load2 <- "uk-englandwales-ndr-2017-listentries-compiled-epoch-0002-baseline-csv.csv"
test0<- read_lines(file2load2, n_max=10)

CorePstCd <- PstCdLSOA.raw %>%
  left_join(EW2, by = c("lsoa11cd" = "ECODE")) %>%
  select(Postcode, lsoa11cd, LAD11CD, LAD11NM,Admin_ward_code = Admin_ward_code.x, WD16NM)

setwd(file.path(basewd, "VOA data"))

CompanyData <- read_delim(file2load2, delim = "*", col_names = F) %>%
  mutate(X15 = gsub(" ", "", X15)) %>%
  left_join(., CorePstCd, by = c("X15" = "Postcode"))

#X10 address
#X11 is road
#X15 is postcode

LSOABiz <- CompanyData %>%
  filter(!is.na(LAD11CD)) %>%
  mutate(X6 = gsub("( AND PREMISES)|( & PREMISES)", "",X6),
         X6 = gsub("(AND PREMISES)|(& PREMISES)", "",X6)) %>%
  group_by(lsoa11cd, X6) %>%
  summarise(counts = n(),
            LAD11CD = first(LAD11CD),
            LAD11NM = first(LAD11NM),
            Admin_ward_code = first(Admin_ward_code),
            WD16NM = first(WD16NM))

FirmCounts <- LSOABiz %>%
  group_by(X6) %>%
  summarise(counts = n()) %>%
  arrange(-counts)

Dom <- FirmCounts %>%
  filter(grepl("DOMESTIC",X6))

HOLS <- FirmCounts %>%
  filter(grepl("HOLIDAY",X6))

Catering <- FirmCounts %>%
  filter(grepl("SELF CATER",X6))

```

Look at Westminster

```{r}

West <- LSOABiz %>%
  filter(LAD11CD=="E09000033")

Catering <- West %>%
  filter(grepl("SELF CATER",X6))

test <- list(HammersmithLONDATA, WestminsterLONDATA, ChelseaLONDATA, CoLLONDATA, CamdenLONDATA, LambethLONDATA) %>% 
  map_df(~.x) %>%
  group_by(LAD11CD) %>%
  summarise(
    LAD11NM =first(LAD11NM),
    LowUse = sum(LowUse, na.rm= T),
    Homes = sum(Homes, na.rm= T),
    Pop = sum(Pop, na.rm= T)
            ) %>%
  filter(Homes >2000) %>%
  mutate(LowUseH = LowUse/Homes,
         LowUseP = LowUse/Pop)

```

```{r}

SalesLAD <- prices %>%
  group_by(Admin_district_code) %>%
    filter(X5 %in% c("D", "S", "T", "F")) %>%
    summarise(sales = n(), 
            mean = mean(X2, na.tm = TRUE))

SalesWard <- prices %>%
  group_by(Admin_ward_code) %>%
    filter(X5 %in% c("D", "S", "T", "F")) %>%
    summarise(sales = n(), 
            mean = mean(X2, na.tm = TRUE))

SalesWard %>% ggplot(aes(x = sales, y = mean )) + geom_point()

```

```{r}
test <- ChelseaLONDATA %>%
            # select(WardLowUsePerc, WardLowUse) %>%
             mutate( WardLowUsePerc = ifelse(is.na(WardLowUsePerc),0, WardLowUsePerc)) %>%
             summarise( maxPerc = max(WardLowUsePerc, na.rm=TRUE),
                        minPerc = min(WardLowUsePerc, na.rm=TRUE),
                        diffPerc = maxPerc-minPerc,
                        max = max(WardLowUse, na.rm=TRUE),
                        min = min(WardLowUse, na.rm=TRUE),
                        diff = max-min,
                        mean = mean(WardLowUse, na.rm=TRUE),
                        median = median(WardLowUse, na.rm=TRUE),
                        LAD = sub("DATA", "", "ChelseaLONDATA"),
                        TurnoverPerc = sum(PercTurnover*Homes, na.rm = TRUE)/sum(Homes, na.rm = TRUE),
                        LUPTurnoverPerc = sum(PercTurnover*LowUse, na.rm = TRUE)/sum(LowUse, na.rm = TRUE)
                        )
```

#Match VOA to offshore test work

```{r}

setwd(DataFolder)

Offshore <- read_excel("OV_FULL_2016_02 (with PEP database for lookup).xlsx", sheet = "OV_FULL_2016_02")

test <- Offshore %>%
  filter(!is.na(Price_Paid))

Offshore$Property_Address %>% head(20)

Offshore2 <- Offshore %>%
    select(Property_Address, Postcode) %>%
    mutate(streetAddress = str_split(Offshore$Property_Address, ",", n = 2) %>% 
  map_chr(~.x[1]),
           Postcode = gsub(" ", "",Postcode))

NumberofPostocodes <-CompanyData %>% filter(X15 %in% Offshore2$Postcode) %>%
  group_by(X15) %>%
  summarise(counts = n()) %>%
  left_join(Offshore2, ., by = c("Postcode" = "X15")) %>%
  filter(!is.na(Postcode))

#20,000 flat or apartments
grepl("^flat|^apartment|^penthouse", Offshore2$Property_Address, ignore.case = T) %>% table
ffshore$Property_Address
grepl("suite", Offshore2$Property_Address, ignore.case = T) %>% table
grepl("unit", Offshore2$Property_Address, ignore.case = T) %>% table

test <- Offshore2 %>%
  filter(grepl("unit", Offshore2$Property_Address, ignore.case = T))

#need another regex which looks at number and if followed by apartement etc uses the number as the flat number
Flatregex <- "(the.+)?(flat|apartment|penthouse)(s)?(\\s(no)?(.+?)(\\s|,)(\\s?(to|-)\\s?.+?(\\s|,))?|,)"

Flats <- Offshore2 %>%
  filter(grepl("flat|apartment|penthouse", Property_Address, ignore.case = T))%>%
  mutate(FlatNumber = str_extract(tolower(Property_Address), 
            Flatregex) %>%
           sub("\\s$|,$", "", .)
)

# (the.+)?(flat|apartment|penthouse)(s)?(\\s(.+?)(\\s|,)(\\s?(to|-)\\s?.+?(\\s|,))?|,)
# 
# some flats start with a "the eastern" or "the penthouse" this is optional
# (the.+)?
# 
# the main expression it has to say this! optioanl "s" on the end
# (flat|apartment|penthouse)(s)?
#
#the additional matching characteristics these come either side of and or statement
# ((s)?\\s(.+?)(\\s|,)(\\s?(to|-)\\s?.+?(\\s|,))?|,)
# (A|B)
# 
# 
# A = \\s(.+?)(\\s|,)(\\s?(to|-)\\s?.+?(\\s|,))?
# A = (A1)(A2)
# 
# space followed by some characters followed by a space or a comma this extracts the flat number eg "flat c"
# A1 = \\s(.+?)(\\s|,)
# 
# optional space then either "to" or "-" followed by optional space followed by some characters followed by space or comma. This is to catch things like "flats 1 to 5" or flat 4-6
# A2 = (\\s?(to|-)\\s?.+?(\\s|,))?
# 
# This is simple some are simply called "flat"  or "the penthouse" this catches those
# B = ,

Flats <- Flats%>%
  mutate(leng = str_split(Flats$Property_Address, ",", n = 2) %>% 
  map_chr(~length(.x)) 
  ) %>% 
  filter(leng == 2) %>%
  mutate( address2 = str_split(Property_Address, ",", n = 2) %>% 
  map_chr(~.x[[2]]) %>% 
    gsub("\\s*\\([^\\)]+\\)", "",.)%>% 
    gsub(" and .+", "",.)
  )

Flats2 <- Flats %>%
  group_by(address2) %>%
  summarise(counts = n()) %>%
  arrange(-counts)

test <- Flats2 %>%
  filter(grepl("Dolphin Square", address2, ignore.case = TRUE))

sum(test$counts)

test <- Offshore2 %>%
  filter(grepl("^[a-z]+( [a-z])* house", Offshore2$Property_Address, ignore.case = T))

#car parks
grepl("car", Offshore2$Property_Address, ignore.case = T) %>% table

ParkingSpaces <- Offshore2 %>%
  filter(grepl("^car |^park", Offshore2$Property_Address, ignore.case = T))

#random land
grepl("^land", Offshore2$Property_Address, ignore.case = T) %>% table

test <- Offshore2 %>%
  filter(grepl("^land", Offshore2$Property_Address, ignore.case = T))

Roadnames <- unique(CompanyData$X11) %>% 
  gsub(" north| south| east| west", "",., ignore.case = TRUE) %>% 
  word(.,-1) %>% data_frame(Endings = .) %>%
  group_by(Endings) %>%
  summarise(count = n()) %>%
  arrange(-count) %>%
  mutate(cumsum = cumsum(count),
         cumperc = cumsum/sum(count))

test2 <- test %>%
  group_by(Endings) %>%
  summarise(count = n())

#A list of road names to remove as they cause problems
removeNames <- "park house buildings green end farm estate court bridge hall drove" %>% str_split(., " ", simplify = TRUE)

#match the top most occuring endings to the offshore list and get the counts
TopRoadEndings <- Roadnames$Endings[1:50] %>%  map_dbl(~ sum(grepl(.x, Offshore$Property_Address, ignore.case = TRUE)))

#arrange in order of occurnace and remove the tricky ones
Road2 <- Roadnames[1:50,] %>% mutate(occs = TopRoadEndings) %>%
  arrange(-occs) %>%
  mutate(Endings = tolower(Endings)) %>%
  filter(!(Endings %in% removeNames)) %>%
  mutate(cumsum = cumsum(occs))

length(Offshore$Property_Address)

#need to include thigs like gate

#create the regex string to match
Wordswithapotrophes <- paste0("(( \\w+)('s)?)+ (",paste(c(Road2$Endings),  collapse = "|"),")")
roadmatchregex <- paste0("([0-9]+)", Wordswithapotrophes)

paste0("(( \\w+)('s)?)+(gate,)")

StreetNum <- str_extract(tolower(Offshore2$Property_Address), roadmatchregex )

Offshore3 <- Offshore2 %>% 
  mutate(
    Property_Address =  gsub(" \\(.+?\\)", "", Property_Address),
    class = case_when(
    grepl(Flatregex, Property_Address, ignore.case = T) ~ "domestic",
    grepl("^car |^park", Property_Address, ignore.case = T) ~"parking space",
    grepl("^land", Property_Address, ignore.case = T) ~ "land",
    grepl("^[a-z]+( [a-z])* cottage", Property_Address, ignore.case = T) ~"domestic",
    grepl("(^[0-9]+)(( \\w+)('s)?)+ (court|villas)", Property_Address, ignore.case = T) ~"domestic",
    TRUE ~ "unknown"
  )) %>%
  mutate(Streetadd = StreetNum,
    Streetnumber =gsub("[a-z]", "", StreetNum, ignore.case = T), 
         Street = gsub("[0-9]", "", StreetNum),
    Street = case_when(
      !is.na(Street) ~ Street,
      grepl("^land", Property_Address, ignore.case = T) ~ 
        str_extract(tolower(Property_Address), paste0("(of|at)", Wordswithapotrophes) ) %>%
        gsub("(of |at )", "",.), #begins with land about 13k
      grepl("( \\w+)+(way|gate,)", Property_Address, ignore.case = T) ~ #roadds that er blblablaGATE
        str_extract(tolower(Property_Address), "( \\w+)+(way|gate)," ),
      #Fixing for knightsbridge
       grepl("knightsbridge", Property_Address, ignore.case = T) ~ "knightsbridge",
      grepl("holland", Property_Address, ignore.case = T) ~ "holland",
      TRUE ~ str_extract(tolower(Property_Address), Wordswithapotrophes ) #no street number
    )
    )

test<- Offshore3%>%
  #sorting houses
  mutate(
    HouseName = str_extract(tolower(Property_Address), "^[a-z]+( [a-z])* house")

    )

grep("( \\w+)+(way|gate),", " parkway,")

table(is.na(Offshore3$Street))
table(is.na(Offshore3$Streetnumber))

table(Offshore3$class)

#Everything that is not land about 7000 items 
  test <- Offshore3 %>%
  filter(is.na(Street))  %>%
  filter(grepl(paste0("^[^(land)]"), Property_Address, ignore.case = TRUE))

  #random
    OffshoreHouse <- Offshore3 %>%
  filter(grepl("house", tolower(Property_Address)), class =="unknown")

    #why are there no overlaps?
    #look at needham house

    CompanyHouse <- CompanyData %>%
      mutate(House = tolower(X10)) %>%
  filter(grepl("house", House))

 test <-    CompanyHouse %>%
      filter(grepl( "Darcy House",  House, ignore.case = T))

 #this matchers
   MMhouse <-    OffshoreHouse %>%
      filter(grepl( "Matthew Murray House",  Property_Address, ignore.case = T))

#Darcy matches
Darcy <-    CompanyHouse %>%
      filter(grepl( "Darcy House",  House, ignore.case = T))

#hotel where some rooms are owned offshore
  Needham <-    OffshoreHouse %>%
      filter(grepl( "Needham",  Property_Address, ignore.case = T))

  #All the directorss are registered to the same place and have no digital foot print
  Debenham <-    OffshoreHouse %>%
      filter(grepl( "debenham",  Property_Address, ignore.case = T))

    %>%
    filter(class == "unknown")

table(test$class)

test <- Offshore3 %>% 
  filter(Offshore3$Street %in% tolower(CompanyData$X11)) %>%
  filter(!is.na(Street), class == "unknown") %>%
  group_by(Street) %>%
  summarise(counts = n())

```

#Unique Company's

```{r}

CompCounts <- Offshore %>%
select(c(grep("Proprietor_Name",names(Offshore)) )) %>% unlist %>%
  data.frame(comp = .) %>%
  group_by(comp) %>%
  summarise(counts = n()) %>%
  filter(!is.na(comp))

sum(CompCounts$counts)

plot(density(CompCounts$counts))

test <- CompCounts %>%
  filter(grepl("trustee", comp, ignore.case = T))

```

```{r}

Flats3<- Flats %>% left_join(PstCdLSOA.raw)%>%
   group_by(lsoa11cd) %>%
  summarise(Total = n()) %>%
#  left_join(EW2, by = c("lsoa11cd" = "ECODE")) %>%
  rename(LSOA_CODE = lsoa11cd)

Flats4 <- Flats3 %>% group_by(LAD11NM) %>%
  summarise(Total = sum(Total))

test <-ls(pattern = "LON") %>% 
  map_df(~get(.x))  %>%
   group_by(LSOA_CODE) %>%
   summarise(LowUse = sum(LowUse, na.rm = TRUE)) %>% 
  CalcLowUse() %>%
  select(LSOA_CODE, Homes, LowUse,LAD11NM:MeanPrice) %>%
  left_join(filter(Flats3, !is.na(LSOA_CODE))) %>%
  mutate(Total = ifelse(is.na(Total), 0, Total),
         LowUsePerc = round(Total/Homes,3)*100)

sum(test$Total)

test %>% 
  group_by(LAD11NM) %>%
  summarise(Total = sum(Total)) %>%
  mutate(perc = round(Total/sum(Total),3)) %>%
  arrange(-perc)

g <- MakeLeaflet(test, type = "jyg", bins = c(0, 5,10,15,100))

Dolph<- Offshore %>%
  filter(grepl("dolphin square", Offshore$Property_Address, ignore.case = T))

```

UK has 20k flats owned by offshore companies. 75% are in #London and of those 43% are in #Westminster.

#London Flats owned by offshore companies as a % of all homes. 43% are in #Westminster, 20% in #RBKC

#London has 14k flats owned by offshore companies. 43% are in #Westminster, 20% in #RBKC. Map % of ALL homes

#Address fixing

```{r}

library(poster)
test <- parse_addr(Offshore$Property_Address)

test <- paste(Offshore$Property_Address, "United Kingdom", sep= ", ") %>%
   normalise_addr %>%
  parse_addr

#\([a-z0-9]+\s[a-z0-9]+

    test1 <-paste(Offshore$Property_Address, "United Kingdom", sep= ", ")

    test2 <- paste(Offshore$Property_Address, "United Kingdom", sep= ", ") %>%
     gsub(" \\(.+?\\)", "", .) %>%
            #gsub("(\\s\\()(?=[a-z0-9]+\\s[a-z0-9]+)", ", ", ., ignore.case = TRUE, perl= TRUE) %>%
      gsub("(and )?(car )?(park(ing)?)( space(\\w)?)?", "", .) %>%
    normalise_addr %>%
    parse_addr

    test3 <- test2 %>% mutate(postal_code = gsub(" ", "", Offshore$Postcode)) %>%
      filter(!is.na(postal_code))

    test4 <- test3 %>%
      group_by(postal_code, road) %>%
      summarise(counts = n())

        CompPostRoad <- CompanyData %>%
      group_by(X11, X15) %>%
      summarise(counts = n())

CompAddress <- paste(CompanyData$X8, CompanyData$X15, "United Kingdom", sep =",") %>%
    normalise_addr %>%
    parse_addr

test5 <- CompAddress %>%
      group_by(postal_code, road) %>%
      summarise(counts = n())

  sum(paste(test4$road, test4$postal_code) %in% paste(test5$road, test5$postal_code))

#savved to VOA folder    
#saveRDS(CompAddress, "CompanyAddresseslibpostal.rds")

    test2[1:20] 
    Offshore$Property_Address[1:20]

    sum(grepl("gate", Offshore$Property_Address, ignore.case = TRUE))

test <- Offshore %>%
  select(Property_Address) %>%
  mutate(Property_Address = gsub(" \\(.+?\\)", "", Property_Address) )%>%
  filter(grepl("(and )?(car )?park(ing)? space(s)?|(and )?car park(ing)?( space(s)?)?", Property_Address, ignore.case = T))

test <- Offshore %>%
  select(Property_Address) %>%
  mutate(Property_Address = gsub(" \\(.+?\\)", "", Property_Address) )%>%
  filter(grepl(" bridge", Property_Address, ignore.case = T))

test[1:20,1]

removeNames <- "park house buildings green end farm estate court bridge hall" %>% str_split(., " ", simplify = TRUE)

TopRoadEndings <- Roadnames$Endings[1:50] %>%  map_dbl(~ sum(grepl(.x, Offshore$Property_Address, ignore.case = TRUE)))

Road2 <- Roadnames[1:50,] %>% mutate(occs = TopRoadEndings) %>%
  arrange(-occs) %>%
  mutate(Endings = tolower(Endings)) %>%
  filter(!(Endings %in% removeNames)) %>%
  mutate(cumsum = cumsum(occs))

length(Offshore$Property_Address)

table(is.na(StreetNum))

```