### 2018-07-31 MSOAmodel.Rmd
Get back MSOA model
I made an MSOA model that worked really well. Sadly I crashed the computer and lost all the code. The re-written code produces a rubbish model, I don't know why. Maybe turning it off and on will help...

---
title: "Untitled"
author: "Jonathan Bourne"
date: "31 July 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

Use the MSOA to make a model it has many examples and maps perfectly to the income estimates

Number of MSOA per LAD
```{r}

testMSOA <- EW2 %>%
  group_by(LAD11CD, MSOA11CD) %>%
  summarise(counts = n()) %>%
  group_by(LAD11CD) %>%
  summarise(counts = n())

testWard <- EW2 %>%
  group_by(LAD11CD, Admin_ward_code) %>%
  summarise(counts = n()) %>%
  group_by(LAD11CD) %>%
  summarise(counts = n())

```

```{r}

test <- DistribCompareBootstrapper(ChelseaLONDATA, 1983, 100, type = NULL, PropertyTypes )
test1 <- DistribCompareBootstrapper2(ChelseaLONDATA, 1983, 100, type = NULL, PropertyTypes )

test2 <- DistribCompareBootstrapper2(ChelseaLONDATA, 1983, 100, type = NULL, PropertyTypes, GroupVars = "MSOA11CD" )

```

```{r}
setwd("/home/jonno/Dropbox/SSE/Empty Homes/BootStrapAuthorities2")

#Find what needs to be bootstrapped
BootStrapFiles <- list.files(getwd()) %>% gsub("DATA.rds", "",.)
  newLADsnames <- sub("DATA", "", ls(pattern = "DATA", envir = globalenv())) %>% 
    .[!(. %in% BootStrapFiles)] 
  newLADs <- paste0(newLADsnames, "DATA")

   #only bootstrap if there is something to add.
  #Save each boostrap as you go along to avoid losing it all in a crash
  if(length(newLADsnames)>0){
    newLADs  %>% walk(~{print(.x)
      DistribCompareBootstrapper(get(.x), 1652, 1000, type=NULL, PropertyTypes, GroupVars = "MSOA11CD" )%>%
      saveRDS(., file = paste0(.x, ".rds"))
      }) 
  }

  #once all Bootratps complete load them all into a list
 BootStrapRES <- list.files(getwd()) %>%
   map(~readRDS(.x))

 #name the list
 names(BootStrapRES) <- list.files(getwd()) %>% gsub(".rds", "",.)
```

#Make modelling dataframe

The variables are

mean deprivation decile
Affordability

```{r}

MSOAdep <- LSOADep %>%
  select(LSOA_CODE, Rank = Rank..where.1.is.most.deprived., Decile = Decile..where.1.is.most.deprived.10..of.LSOAs.) %>%
  left_join(select(EW2, MSOA11CD, ECODE), by = c("LSOA_CODE"="ECODE")) %>%
  select(-LSOA_CODE) %>%
  group_by(MSOA11CD) %>%
  summarise_all(mean)

BootDeets <- BootStrapRES %>% 
  map_df(~.x %>%
           mutate(Reference = MedianHomes < MedianLUPs,
                  MeanPropPrice = (HomesValue+LowUseValue)/(Homes+LowUse)) %>%
           select(MSOA11CD, MeanPropPrice, Reference, ID)) %>%
  left_join(select(IncomeEst, MSOA11CD, Yearly.income), by = "MSOA11CD") %>%
  left_join(MSOAdep, by = "MSOA11CD") %>%
  mutate(AffordRatio  = MeanPropPrice/Yearly.income,
         AffordRatio2 = AffordRatio^2,
         AffordRatio3 = AffordRatio^3)

rm(MSOAdep)

#Why do some have 2000?
test <- BootDeets %>%
  group_by(MSOA11CD) %>%
  summarise(Majority = sum(Reference, na.rm = T)/1000,
            Total = n())

BootDeetsMean <- BootDeets %>%
  select(MSOA11CD, Reference, Decile, AffordRatio) %>%
  group_by(MSOA11CD) %>%
  summarise_all(mean) %>%
  mutate(Reference = Reference>0.5)

set.seed(1983)
TestResample <- 1:50 %>% map(~  resample_partition(BootDeetsMean, c(test = 0.1, train = 0.9)))

OpenFormula <- as.formula(Reference~ AffordRatio+ Decile)

BootMod <- MakeBootModels(BootDeets, as.integer(TestResample[[1]]$train))
Predsdf <- MakeBootPreds(BootDeets, BootMod, as.integer(TestResample[[1]]$test))
Accdf   <- MakeBootAccuracy(Predsdf, BootDeets, as.integer(TestResample[[1]]$test))

sum(Accdf$Accuracy>Accdf$AccuracyNull)

mean(Accdf$Accuracy)

CoeffsMega <- 1 %>%
  map_df(~{
    print(.x)
    BootMod <- MakeBootModels(BootDeets, as.integer(TestResample[[.x]]$train), ModForm =  OpenFormula)

    #needs to be mapped again to extract every model not just the first 50!
    Out <- 1:length(BootMod) %>% map_df(~
                          BootMod[[.x]] %>% tidy %>% mutate(ID = .x)               
                                        )%>%
      mutate(Sample = .x)
  })

```