### 2018-08-12 MSOAmodel.Rmd
Simplify Bootstrapping
I refactored the code for bootstrapping and sampling. This made it easier to understand and allowed me to have a single strapping function that can be modified for high or low memory usage.
I also fine tuned the quantiles

---
title: "Untitled"
author: "Jonathan Bourne"
date: "31 July 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

Use the MSOA to make a model it has many examples and maps perfectly to the income estimates

Number of MSOA per LAD
```{r}

testMSOA <- EW2 %>%
  group_by(LAD11CD, MSOA11CD) %>%
  summarise(counts = n()) %>%
  group_by(LAD11CD) %>%
  summarise(counts = n())

testWard <- EW2 %>%
  group_by(LAD11CD, Admin_ward_code) %>%
  summarise(counts = n()) %>%
  group_by(LAD11CD) %>%
  summarise(counts = n())

```

#LAD quartiles are wrong

```{r}

setwd("/home/jonno/Dropbox/SSE/Empty Homes/BootStrapAuthorities2")
test <- DistribCompareBootstrapper(ChelseaLONDATA, 1983, 501, type = NULL, PropertyTypes, GroupVars = "class"  )

ChelTest1 <- DistribCompareBootstrapper(ChelseaLONDATA, 1983, 501, type = NULL, PropertyTypes, GroupVars = "MSOA11CD" )

test2 <- DistribCompareBootstrapper(ChelseaLONDATA, 1983, 501, type = NULL, PropertyTypes, GroupVars = c("MSOA11CD", "class"))

MSOAList <- unique(test2$MSOA11CD) %>%
  map(~{
test2 %>%
  filter(MSOA11CD == .x)
})

test3 <- MSOAList[[21]] %>%
  CleanForPlotClassDiff

MSOAList[[21]] %>%
CleanForPlotClassDiff()  %>%
PlotClassDiff(., "Kensington and Chelsea by MSOA")

test %>%
  mutate(LowUse = Homes) %>%
  CleanForPlotClassDiff()  %>%
  PlotClassDiff(., "Kensington and Chelsea full borough")

test %>%
#  mutate(LowUse = Homes) %>%
  CleanForPlotClassDiff2()  %>%
  PlotClassDiff(., "Kensington and Chelsea full borough")

CleanForPlotClassDiff2 <- function(df){ #relative to the homes distrib
  df %>% 
  group_by(ID, class) %>%
  summarise(Homes = sum(Homes), LowUse = sum(LowUse)) %>%
  group_by(ID)%>%
  mutate(RatioExvsAct = LowUse/sum(LowUse),
         HomesRatio = Homes/sum(Homes),
         RatioExvsAct = (RatioExvsAct/HomesRatio)-1)
}

```

```{r}

PropertyTypes <- c("D", "S", "T", "F")
setwd("/home/jonno/Dropbox/SSE/Empty Homes/BootstrapMSOA")

#Find what needs to be bootstrapped
BootStrapFiles <- list.files(getwd()) %>% gsub("DATA.rds", "",.)
  newLADsnames <- sub("DATA", "", ls(pattern = "DATA", envir = globalenv())) %>% 
    .[!(. %in% BootStrapFiles)] 
  newLADs <- paste0(newLADsnames, "DATA")

  test <- ls(pattern = "DATA") %>% map_df(~
 data.frame(LAD = .x, LSOA = get(.x) %>% nrow(),Homes = sum(get(.x)$Homes, na.rm = T))) %>%
    arrange(Homes)

  newLADs <- newLADs[match(test$LAD, newLADs)] %>% .[!is.na(.)]

  #only bootstrap if there is something to add.
  #Save each boostrap as you go along to avoid losing it all in a crash
  if(length(newLADsnames)>0){
    newLADs  %>% walk(~{print(.x)
      DistribCompareBootstrapper(get(.x), 1652, 501, type=NULL, PropertyTypes, GroupVars = "MSOA11CD", Limit =1.3e5  )%>%
      saveRDS(., file = paste0(.x, ".rds"))
      gc()
      }) 
  }

#once all Bootratps complete load them all into a list
 BootStrapRES <- list.files(getwd()) %>%
   map(~readRDS(.x))

 #name the list
 names(BootStrapRES) <- list.files(getwd()) %>% gsub(".rds", "",.)

```

```{r}

library(Rtsne)

# dfmat <- BootDeetsMean %>%
#   select(AffordDiff, Decile, AffordRatio, AffordRatio2, DecileDiff) %>%
#   as.matrix()

dfmat <- ModelMSOAdf %>%
  select( Tourism, AffordRatio, AffordRatio2) %>%
  as.matrix()

set.seed(12358)
TsnecorOut1 <-dfmat  %>%
  Rtsne(verbose = TRUE, eta = 1, max_iter = 5000, perplexity = 50, check_duplicates = FALSE)

# tsne_corplot <- data.frame(x1 = TsnecorOut1$Y[,1],
#                         y1 = TsnecorOut1$Y[,2]) %>%
#   bind_cols(BootDeetsMean) %>%
#   mutate(MedianCheck = MSOAHomesMedian<MSOALowUseMedian)

tsne_corplot <- data.frame(x1 = TsnecorOut1$Y[,1],
                        y1 = TsnecorOut1$Y[,2]) %>%
  bind_cols(ModelMSOAdf)

tsne_corplot %>%
    ggplot() +
  geom_point(aes(x=x1, y=y1, color= TourDepCat==1|TourDepCat==2))

BootDeetsMean %>%
  ggplot(aes(x = LowUseConc, y = AffordDiff, colour = Reference)) + geom_point()

BootDeetsMean %>%
  filter(sales <1) %>%
  ggplot(aes(x = sales, y = LowUse, colour = Reference)) + geom_point()

```

#Sample quantiles

```{r}

#Create a data frame of LSO
PriceCounts <-prices %>%
  filter(X5 %in% c("D", "S", "T", "F")) %>%
  group_by(LSOA11CD) %>%
  summarise(counts = n()) %>%
  arrange(counts)
##The subsampling is 1/10 of the number of homes in each LSOA, this stops crashes
PriceCounts<- EW2 %>%
  select(LSOA11CD = ECODE, Homes, Pop, LAD11CD, MSOA11CD) %>%
  left_join(.,PriceCounts, by = "LSOA11CD") %>%
  mutate(counts = ifelse(is.na(counts), 0, counts),
         Homes2 = Homes/10) %>%
  filter(counts!=0)

#Create a list of price vectors by LSOA
#This process can take a while
setwd(basewd)
if(file.exists("LSOAPriceList2.rds")){

  LSOAPriceList <- readRDS("LSOAPriceList2.rds")

}else{

  LSOAPriceList <- 1:nrow(PriceCounts) %>% map(~{
   print(.x)
    prices %>%
      filter(LSOA11CD==PriceCounts$LSOA11CD[.x], 
             X5 %in% c("D", "S", "T", "F")) %>%
      pull(X2)
    })
  names(LSOAPriceList) <- PriceCounts$LSOA11CD
  saveRDS(LSOAPriceList, "LSOAPriceList2.rds")

}

if(file.exists("AllDataQuartiles.rds")){

  AllDataQuartiles <- readRDS("AllDataQuartiles.rds")

}else{
  #Subset list to only the LADs used in analyis
  LSOAPriceList2 <- LSOAPriceList[names(LSOAPriceList) %in% unique(DATAdf$LSOA11CD)]
  PriceCounts2 <- PriceCounts %>%
    filter(LSOA11CD %in% unique(DATAdf$LSOA11CD))

  AllDataQuartiles <- StratifiedBoot(LSOAPriceList2, PriceCounts2$LSOA11CD, PriceCounts2$Homes2, samples = 501)

  saveRDS(AllDataQuartiles, "AllDataQuartiles.rds")

  rm(list = c("LSOAPriceList2", "PriceCounts2"))

}

rm(LSOAPriceList)

AllDataQuartilesSummary <- AllDataQuartiles %>%
  summarise_all(funs(mean, sd))

BasicQuartiles <- prices$X2 %>% .[prices$LAD11CD %in% unique(DATAdf$LAD11CD)] %>% quantile(.)

#difference between the two up to 8% this has knock on effects later in the analysis if the quartiles are not calculated correctly
AllDataQuartilesSummary[2:4]/BasicQuartiles[2:4]

prices <- prices %>%
  mutate(CountryClass = cut(X2, AllDataQuartilesSummary[2:4] %>% c(0,., Inf), 
                              labels =     c("Lower", "Lower-Mid", "Upper-Mid", "Upper"), 
                              right = F) %>% fct_relevel(., "Upper", after = 3))

rm(BasicQuartiles);rm(AllDataQuartiles);rm(AllDataQuartilesSummary)

```

#Calc Real Quantiles for all data

This takes ages and gives an almost identical number to simple running quantiles on the whole dataset

```{r Real Quantile Chunk}

test <-  AllDataQuartiles %>%
   select(-X0.,-X100.) %>%
   summarise_all(funs(mean, sd)) %>%
   mutate(X25Perc = X25._sd/X25._mean,
          X50Perc = X50._sd/X50._mean,
          X75Perc = X75._sd/X75._mean)

prices %>%
   filter(X5 %in% c("D", "S", "T", "F")) %>%
  pull(X2) %>% quantile()

```

#MSOA Quartiles

I'm not sure this is useful

```{r}
#Create a list of vectors where each element of the list is an MSOA and the vector elements are its LSOA
 LSOAinMSOA <- unique(PriceCounts$MSOA11CD) %>% map(~{
   print(.x)
    PriceCounts %>%
      filter(MSOA11CD==.x) %>%
      pull(LSOA11CD)
 })
 #name the elements of the list
 names(LSOAinMSOA) <- unique(PriceCounts$MSOA11CD)

 #Create a bootstrap for each of the MSOA

 MSOABootQuartiles <- unique(PriceCounts$MSOA11CD) %>% map_df(~{
   print(.x)
   PriceCounts2 <- PriceCounts %>%
     filter(MSOA11CD == .x)

   LSOAPriceList2 <- LSOAPriceList[names(LSOAPriceList) %in% PriceCounts2$LSOA11CD]

   StratifiedBoot(LSOAPriceList2, PriceCounts2$LSOA11CD, PriceCounts2$Homes2, samples = 100) %>%
     mutate(MSOA11CD=.x)

 })

 saveRDS(MSOABootQuartiles, "MSOABootQuartiles.rds")

 MSOABootQuartiles2 <- MSOABootQuartiles %>%
   select(-X0.,-X100.) %>%
   group_by(MSOA11CD) %>%
   summarise_all(funs(mean, sd)) %>%
   mutate(X25Perc = X25._sd/X25._mean,
          X50Perc = X50._sd/X50._mean,
          X75Perc = X75._sd/X75._mean)  %>%
   left_join(., EW2%>%
               select(MSOA11CD, LAD11CD, LAD11NM) %>% distinct()) %>%
   left_join(., IncomeEst) %>%
   select(MedianPrice=X50._mean, Yearly.income, MSOA11CD, LAD11CD, LAD11NM) %>%
   mutate(AffordRatio = MedianPrice/Yearly.income,
          AffordRank = percent_rank(AffordRatio))
```

#Local value

Here I create a graph of the MSOA. 
I calculate the avereage distance by LAD.
I then use that distance to find all the MSOA within that jump distance.
I then calculate the mean and quartiles of the local area for each MSOA
From there I can calculate the local area affordability and thus the relative local area affordability

```{r}

 MSOAAdjacency <- readShapeSpatial(file.path("/home/jonno/Dropbox/SSE/Empty Homes/ShapeFiles/Middle_Layer_Super_Output_Areas_December_2011_Super_Generalised_Clipped_Boundaries_in_England_and_Wales",
                                             "Middle_Layer_Super_Output_Areas_December_2011_Super_Generalised_Clipped_Boundaries_in_England_and_Wales.shp"))

 test <- poly2nb(MSOAAdjacency)

 MSOAAdjMat = nb2mat(test, style="B",zero.policy=T)

#Add MSOA codes so that they are added as the vertex names
colnames(MSOAAdjMat) <- MSOAAdjacency$msoa11cd

MSOAgraph <- graph.adjacency(MSOAAdjMat, mode = "undirected")

#Add in Lad

MSOAgraph <- set_vertex_attr(MSOAgraph, 
                name = "LAD11CD", 
                value = MSOAtoLAD$LAD11CD[match(get.vertex.attribute(MSOAgraph, "name"),
                                          MSOAtoLAD$MSOA11CD)]
                )

#Subset by LAD into a list of graphs

Subgraph <- function(g, KeepNodes){

  VertexNames <-get.vertex.attribute(g, "name")
  VertexNames<- VertexNames[!(VertexNames %in% KeepNodes)]

g2 <- delete_vertices(g, VertexNames)

g2
}

AllLADsGraph <- unique(MSOAtoLAD$LAD11CD) %>% map(~{
  KeepNodes <-MSOAtoLAD %>%
  filter(LAD11CD ==.x)
KeepNodes2 <- KeepNodes$MSOA11CD

  Subgraph(MSOAgraph, KeepNodes2)
})

#name the elements by the LAD represented
names(AllLADsGraph) <- unique(MSOAtoLAD$LAD11CD)

#calculate mean distance and diameter of each lad
MSOAconnections <- AllLADsGraph %>% map_df(~{

  tibble(distance = mean_distance(.x), diameter = diameter(.x), MSOA = vcount(.x))

}) %>%
  mutate(LAD11CD = names(AllLADsGraph)) %>%
  filter(MSOA !=1) #removes city of london which only has a single MSOA and so can't be included in the calc

MSOAconnections %>%
  ggplot(aes(x=diameter )) + geom_density()

#Using mean distance is best
summary(MSOAconnections$distance)
summary(MSOAconnections$diameter)

# find names of local MSOA using an x jump distance
EgoSize <- tibble(ego_size = ego_size(MSOAgraph, 2), MSOA11CD = get.vertex.attribute(MSOAgraph, "name")) %>%
  left_join(MSOAtoLAD)
##Ego size of jump two has a mean almost identical to the mean number of MSOA in a LAD. However the Ego size is less skewed.
mean(EgoSize$ego_size)

MSOAtoLAD %>%
  group_by(LAD11CD) %>%
  summarise(counts = n(),
            LAD11NM = first(LAD11NM)) %>%
  pull(counts) %>% mean

test <-  tibble(MSOA11CD = ego(MSOAgraph, 1, nodes = "E02000001")[[1]] %>% names) %>%
  left_join(MSOAtoLAD)

##names of the MSOA in each area.
MSOAEgoList <- ego(MSOAgraph, 2) %>%map(~names(.x))
names(MSOAEgoList) <- get.vertex.attribute(MSOAgraph, "name")

```

```{r}
#Create a list of vectors where each element of the list is an MSOA and the vector elements are its LSOA
 LSOAinMSOAEgo <- MSOAEgoList %>% map(~{
   #print(.x)
    PriceCounts %>%
      filter(MSOA11CD %in% .x) %>%
      pull(LSOA11CD)
 })

 ##name the elements of the list
 names(LSOAinMSOAEgo) <- names(MSOAEgoList)

 ##Create a bootstrap for the ego of each MSOA
 NeighbourhoodQuartiles <- names(MSOAEgoList) %>% map_df(~{

   print(.x)
   target <- MSOAEgoList[[.x]]
   PriceCounts2 <- PriceCounts %>%
     filter(MSOA11CD %in% target)

   LSOAPriceList2 <- LSOAPriceList[names(LSOAPriceList) %in% PriceCounts2$LSOA11CD]

   Out <-StratifiedBoot(LSOAPriceList2, PriceCounts2$LSOA11CD, PriceCounts2$Homes2, samples = 10) %>%
     mutate(MSOA11CD=.x)

   Out

 })

```

Do westminster and lambeth fall into each others local area?
need to check and make sure they do!
```{r}
#Calculate mean local income
LocalIncome <- names(MSOAEgoList) %>% map(~{
  print(.x)
#.x <- "E02000001"
Group <-MSOAEgoList[[.x]]

Incomest2 <- IncomeEst %>%
  left_join(select(MSOAtoLAD, MSOA11CD, Homes),  by="MSOA11CD")

Out <- Incomest2 %>%
  filter(MSOA11CD %in% .x) %>%
  summarise(LocalIncome =sum(Yearly.income* Homes)/sum(Homes)) %>% pull()

Out
}) %>% unlist %>%
  tibble(MSOA11CD = names(MSOAEgoList), LocalIncome = .)

LocalAfford <-NeighbourhoodQuartiles %>%
  select(-ID,-X0., -X100.) %>%
    group_by(MSOA11CD) %>%
  summarise_all(funs(mean, sd, min, max)) %>%
  mutate(perc = X50._sd/X50._mean) %>%
  left_join(LocalIncome, by = "MSOA11CD") %>%
  mutate(LocalAfford = X50._mean/LocalIncome) %>%
  select(MSOA11CD, MedianLocalHomesPrice = X50._mean, LocalAfford)

```

#Complex models

These models work at MSOA level only and include data from the LAD level
Tehy are rubbish so aren't included in the main work. Just kept here so I don't forget I have already checked.

```{r}

DATAdf %>%
  arrange(LAD11NM) %>% pull(LAD11NM) %>% unique

MSOAMean2 <-MSOAMean %>%
  mutate(RelativeAfford = AffordRatio/LADAffordRatio,
         RelativeTourismDens = TourismDensity/LADTourismDensity,
         RelativeTourismDens = ifelse(is.finite(RelativeTourismDens), RelativeTourismDens, 100)) %>%
  filter(!is.na(HighVal)) #Temporary  until I work out how to get rid of those weird MSOA

KitchenSink <- as.formula(Reference ~ TourismDensity+ AffordRatio+ AffordRatio2+ RelativeAfford +RelativeTourismDens+
                            LADTourismDensity+ LADAffordRatio+ LADAffordRatio2)

KitchenSink2 <- as.formula(Reference ~RelativeAfford +RelativeTourismDens+
                            LADTourismDensity+ LADAffordRatio)
Formulas2 <- c(as.formula(Reference ~ TourismDensity+ AffordRatio+ AffordRatio2),
               KitchenSink)

set.seed(1983)
TestResample <- 1:100 %>% map(~  resample_partition(MSOAMean2 , c(test = 0.2, train = 0.8)))

MSOAModels2 <-LinearClassifierCombiner(MSOAMean2,  Outcomes, Formulas2, TestResample) %>%
  mutate(IncAfford2 = grepl("2", Formula),
           LAD = grepl("LAD", Formula)
         )

#No difference
MSOAModels2 %>%
  group_by(Dependent, IncAfford2, LAD) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            ) %>%
  ungroup %>%
   mutate(Dependent = ifelse(Dependent=="HighLUP", "High LUP %", "High LUP value"))

library(ranger)

#Just under 78% better than all other models
ResOut <- MSOAMean2 %>%
  mutate(Reference = HighLUP) %>% 
  TreeResampledModelPerf(., TestResample, KitchenSink)

ResOut2 <- MSOAMean2 %>%
  mutate(Reference = HighLUP) %>%
  TreeResampledModelPerf(., TestResample, KitchenSink2)

ResOut %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )

ResOut2 %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )

```

```{r}
#This model uses information from the over all LUP as a result I need to subset the data by lups then train on the MSOA
set.seed(1983)
TestResample <- 1:100 %>% map(~  resample_partition(LADMean , c(test = 0.2, train = 0.8)))

ComplexModels <- c("BothHighLUP", "BothHighVal", "MSOALUPLADVal", "LADHighLUP", "HighLUP", "LADHighVal") %>%
  map_df(~{
    print(.x)
    ResOut2 <- MSOAMean2 %>%
      rename_(Reference = .x) %>%
      TreeResampledModelPerf(., TestResample, KitchenSink, CombineData = TRUE) %>%
      mutate(Dependent = .x)
})

ComplexModels %>%
  group_by(Dependent) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )

KitchenSink2 <- as.formula(Reference ~RelativeAfford +RelativeTourismDens+
                            LADTourismDensity+ LADAffordRatio)

ComplexModels2 <- c("BothHighLUP", "BothHighVal", "MSOALUPLADVal", "LADHighLUP", "HighLUP", "LADHighVal") %>%
  map_df(~{
    print(.x)
    ResOut2 <- MSOAMean2 %>%
      rename_(Reference = .x) %>%
      TreeResampledModelPerf(., TestResample, KitchenSink2, CombineData = TRUE) %>%
      mutate(Dependent = .x)
})

ComplexModels2 %>%
  group_by(Dependent) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )

Modeldf <-MSOAMean2 %>%
      rename_(Reference = "LADHighLUP")

```

#Collaspy tree voting
```{r}

test <- MSOAMean2 %>%
      rename_(Reference = "LADHighLUP") %>%
MSOATreeVoting(., TestResample, KitchenSink )

test %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )

```