### 2018-08-12 PaperCode.Rmd

---
title: "Untitled"
author: "Jonathan Bourne"
date: "9 August 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
Mang en natt har jeg sittet opp og skrevet dype romaner paa dette skrivebordet. Det er pent brukt, men har gitt meg graa haar. Det er som min Moby Dick, jeg har aldri helt klart aa temme det, boken er halvskrevet, hodet fult, hjertet tomt. Derfor maa det bort, kona orker ikke at jeg sitter oppe og stirrer paa uskrevne linjer i min moleskinen mer.

```{r}
SubCode <- "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode"
setwd(SubCode)
source("Setup.R")

setwd(Functions)
list.files() %>% map(source)

Figures <- "/home/jonno/Dropbox/Apps/ShareLaTeX/Empty Homes Write up 2/Figures" #file.path(basewd, "Figures")
TexTables <- "/home/jonno/Dropbox/Apps/ShareLaTeX/Empty Homes Write up 2/Tables"
suppressMessages(source(file.path(CommonCode, "AuxdataLoad.R")))
```

#Process all areas

This chunk processes all the areas and combines into a single dataframe
It also cleans all postcodes that are in other LADs out... should it just combine them into the other LADs LSOA?
```{r}
#creates a data frame called DATAdf
suppressMessages(source(file.path(SubCode, "LOADandProcessLADData.R")))

rm(PstCdLSOA.raw)
```

#MSOA to LAD
```{r}
MSOAtoLAD <- EW2 %>%
  group_by(MSOA11CD, LAD11CD, LAD11NM) %>%
  summarise(TotMSOA = n(),
         Homes = sum(Homes),
         Region = first(Region)) %>%
  arrange(-TotMSOA) %>%
  group_by(MSOA11CD) %>%
  mutate(counts = n()) %>%
  summarise(LAD11CD = first(LAD11CD),
            TotMSOA = first(TotMSOA),
            counts = first(counts),
            LAD11NM = first(LAD11NM),
            Homes = first(Homes),
            Region = first(Region))
```

#Get Population Statistics

This chunk calculates how much data cover I have using several metrics

```{r}
DataCover <-DATAdf  %>%
                #filter(!is.na(LAD11NM)) %>%
  summarise(
           'LADs' = length(unique(DATAdf$LAD11CD)), 
           '% of all LADs' = length(unique(DATAdf$LAD11CD))/length(unique(EW2$LAD11CD))*100,
           LUPs = sum(LowUse, na.rm = T),
           Homes = sum(Homes, na.rm = T),
           '% of all Homes' = Homes/sum(EW2$Homes)*100,
           Population = sum(Pop, na.rm= T),
           '% of total population' = Population/sum(EW2$Pop)*100) %>%
  mutate_if(.<1, funs(signif(.,3))) %>%
  mutate_if(.>1, as.integer) %>%
  mutate_all(as.character) %>% 
  gather

DataCover

xtable(DataCover, caption="Summary of collected data coverage",
       label = "tab:DataCover") %>% 
  print(., type="latex", file=file.path(TexTables,"DataCover.tex"))

rm(DataCover)

```

#Calculate Quantiles

This rather long winded way of getting the quartiles is important as it weights the sampling by Homes in an area. Relative proportion of sales in an area is not the same as homes. Some areas have higher turnover other lower. This means the quartiles need to be repeatedly sampled to get the right quartiles for the whole dataset.

In the below chunk I have sampled the data 501 times, however it is only necessary to sample one the quartiles don't change.

```{r}

#Create a data frame of LSO
PriceCounts <-prices %>%
  filter(X5 %in% c("D", "S", "T", "F")) %>%
  group_by(LSOA11CD) %>%
  summarise(counts = n()) %>%
  arrange(counts)
##The subsampling is 1/10 of the number of homes in each LSOA, this stops crashes
PriceCounts<- EW2 %>%
  select(LSOA11CD = ECODE, Homes, Pop, LAD11CD, MSOA11CD) %>%
  left_join(.,PriceCounts, by = "LSOA11CD") %>%
  mutate(counts = ifelse(is.na(counts), 0, counts),
         Homes2 = Homes/10) %>%
  filter(counts!=0)

#Create a list of price vectors by LSOA
#This process can take a while
setwd(basewd)
if(file.exists("LSOAPriceList2.rds")){

  LSOAPriceList <- readRDS("LSOAPriceList2.rds")

}else{

  LSOAPriceList <- 1:nrow(PriceCounts) %>% map(~{
   print(.x)
    prices %>%
      filter(LSOA11CD==PriceCounts$LSOA11CD[.x], 
             X5 %in% c("D", "S", "T", "F")) %>%
      pull(X2)
    })
  names(LSOAPriceList) <- PriceCounts$LSOA11CD
  saveRDS(LSOAPriceList, "LSOAPriceList2.rds")

}

if(file.exists("AllDataQuartiles.rds")){

  AllDataQuartiles <- readRDS("AllDataQuartiles.rds")

}else{
  #Subset list to only the LADs used in analyis
  LSOAPriceList2 <- LSOAPriceList[names(LSOAPriceList) %in% unique(DATAdf$LSOA11CD)]
  PriceCounts2 <- PriceCounts %>%
    filter(LSOA11CD %in% unique(DATAdf$LSOA11CD))

  AllDataQuartiles <- StratifiedBoot(LSOAPriceList2, 
                                     PriceCounts2$LSOA11CD, 
                                     PriceCounts2$Homes2, 
                                     samples = 501) #change to 1 for a quicker solution has no effect on outcome

  saveRDS(AllDataQuartiles, "AllDataQuartiles.rds")

  rm(list = c("LSOAPriceList2", "PriceCounts2"))

}

AllDataQuartilesSummary <- AllDataQuartiles %>%
  summarise_all(funs(mean, sd))

BasicQuartiles <- prices$X2 %>% .[prices$LAD11CD %in% unique(DATAdf$LAD11CD)] %>% quantile(.)

#difference between the two up to 8% this has knock on effects later in the analysis if the quartiles are not calculated correctly
AllDataQuartilesSummary[2:4]/BasicQuartiles[2:4]

#Add the balanced quartiles on to the prices.
prices <- prices %>%
  mutate(CountryClass = cut(X2, AllDataQuartilesSummary[2:4] %>% c(0,., Inf), 
                              labels =     c("Lower", "Lower-Mid", "Upper-Mid", "Upper"), 
                              right = F) %>% fct_relevel(., "Upper", after = 3))

rm(BasicQuartiles);rm(AllDataQuartiles);rm(AllDataQuartilesSummary);rm(LSOAPriceList)

```

#BootStrap

This section bootstraps the price classes for each property then aggregates it by quartiles per LAD
From this we can find out if the majority of a LADs Low Use properties are investment or not by comparing the median values.

The same can be done at MSOA level but the areas are so small that you don't get enough variation in the house price levels to get a meaningful result. A better way to do this would be to use the ego network of MSOA around a target to find the "Neighbourhood LAD" and see if that is investment. However as I don't have the the low use property data for all adjoining MSOA this is not possible.

##Quartiles Bootstrap

Needed for creating the skew plots
```{r}
setwd(file.path(basewd, "BootstrapQuartiles"))
BootstrapAllData(DATAdf, LimitValue = 1.4e5, Reps = 501)
BootStrapQuartiles <- LoadBootstrapData()

```

##MSOA Bootstrap
Used for most of the analysis
```{r}
setwd(file.path(basewd, "BootstrapMSOA"))
BootstrapAllData(DATAdf, LimitValue = 1.4e5, Reps = 501, GroupingVars = "MSOA11CD")
BootStrapMSOA <- LoadBootstrapData()

```

#Low Use quartile plot

This chunk plots which of the LADs affordability quartiles the LUPs fall into.

```{r}
CleanForPlotClassDiff <- function(df){ #not a seperate function for ease of editing
  df %>%  
  group_by(ID, Class) %>%
  summarise(Homes = sum(Homes), 
            LowUse = sum(LowUse)) %>%
  group_by(ID)%>%
  mutate(LowUseDistrib = LowUse/sum(LowUse),
         HomesDistrib = Homes/sum(Homes)) %>%
  mutate(LowUseRatio1 = (LowUseDistrib/0.25)-1,
         HomesRatio1 = (HomesDistrib/0.25)-1,
         LowUseRatio2 = (LowUseDistrib-HomesDistrib)/HomesDistrib) %>%
    ungroup
}

LADS<-DATAdf%>%
  group_by(LAD11CD) %>%
  summarise(LAD11NM = first(LAD11NM)) %>%
  arrange(LAD11NM)

#Plot for all data
All_data_class <- BootStrapQuartiles %>% 
  bind_rows %>%
  select(CountryClass, ID, LowUse, Homes) %>%
  rename(Class = CountryClass)

BootStrapQuartiles$E09000020  %>% CleanForPlotClassDiff()  %>%PlotClassDiff(Var = "HomesRatio1")
BootStrapQuartiles$E08000032 %>% CleanForPlotClassDiff()  %>%PlotClassDiff(Var = "HomesRatio1")
All_data_class %>% CleanForPlotClassDiff()  %>%PlotClassDiff(Var = "LowUseRatio2")

#All data summary
All_data_class %>% 
  group_by(Class,ID) %>%
  summarise_all(sum) %>%
  group_by(Class) %>%
  summarise_all(mean) %>%
  CleanForPlotClassDiff()

#Create multi panel plot for two areas and all data
bind_rows(
CleanForPlotClassDiff(BootStrapQuartiles$E09000020) %>% mutate(Area = "Kensington and Chelsea"),
CleanForPlotClassDiff(BootStrapQuartiles$E08000032) %>% mutate(Area = "Bradford"),
CleanForPlotClassDiff(All_data_class) %>% mutate(Area = "All Data")
) %>%
  mutate(Area = factor(Area, levels = c("Bradford", "All Data", "Kensington and Chelsea"))) %>%
  PlotClassDiff(Var = "LowUseRatio1") +
    facet_wrap(.~Area) +
  ggtitle("Difference in Low-Use Property distribution by Price Quartile")
SaveFig("ClassdiffAllData.pdf")

rm(BootStrapQuartiles);rm(All_data_class)
```

#Tourism data
This section Loads and processes tourism Data

```{r}
setwd(file.path(basewd, "VOA data"))

#Creates Tourism dataframe at LSOA level
Tourism <- read_delim("uk-englandwales-ndr-2017-listentries-compiled-epoch-0002-baseline-csv.csv", 
                          delim = "*", col_names = F) %>%
  mutate(X15 = gsub(" ", "", X15)) %>%
  left_join(., CorePstCd, by = c("X15" = "Postcode")) %>%
  filter(!is.na(LAD11CD)) %>%
  filter(grepl("(SELF CATER)|(HOLIDAY)|(HOTEL)|(HOSTEL)|(GUEST)",X6))  %>%
  mutate(Guest = grepl("(SELF CATER)|(HOLIDAY)|(GUEST)",X6),
         Hotel = grepl("(HOTEL)|(HOSTEL)",X6)) %>%
   group_by(LSOA11CD, X6) %>%
  summarise(Tourism = n(),
            LAD11CD = first(LAD11CD),
            LAD11NM = first(LAD11NM),
            MSOA11CD = first(MSOA11CD),
            Pop = first(Pop),
            Guest = sum(Guest),
            Hotel = sum(Hotel))

```

#Create LAD and MSOA DATA frames

This chunk creates useful data frames for the rest of the analysis

##LAD Data

```{r}
LADModelData <-CreateGeogModelData("LAD", file.path(basewd, "BootstrapMSOA"))
LADMean <- MeanModelData(LADModelData, "LAD") %>%
  mutate(AffordRank2 = scale(AffordRank)^2,
           AffordRatio2 = scale(AffordRatio)^2,
         TourismDensityRank = rank(TourismDensity, ties.method = "max"),
         LowUseType =case_when(
    "TRUE"==HighLUP & "TRUE" ==HighVal ~ 1, # All high
    "TRUE" == HighLUP & "FALSE" ==HighVal ~ 2, #High LUP in MSOA but not in LAD
    "FALSE"==HighLUP & "TRUE" ==HighVal ~ 3, #High LUP in LAD but not MSOA
    "FALSE"==HighLUP & "FALSE"==HighVal ~ 4 #LOW LUP in both LAD and MSOA
  ) %>% as.factor,
  LowUseAfford =case_when(
    "TRUE"==HighLUP & AffordRank>=0.5 ~ 1, # All high
    "TRUE" == HighLUP & AffordRank<0.5 ~ 2, #High LUP in MSOA but not in LAD
    "FALSE"==HighLUP & AffordRank>=0.5 ~ 3, #High LUP in LAD but not MSOA
    "FALSE"==HighLUP & AffordRank<0.5 ~ 4 #LOW LUP in both LAD and MSOA
  ) %>% as.factor)  %>%
  left_join(MSOAtoLAD %>%
              group_by(LAD11CD) %>%
              summarise(LAD11NM = first(LAD11NM)), by = "LAD11CD")

  LADMean %>%
  ggplot(aes(x= AffordRank, y = LowUseRank, colour = LowUseAfford)) + geom_point() +
    geom_vline(xintercept  = 0.5)+
    geom_hline(yintercept = 0.5) +
    labs(title = "Seperating Low-Use type  by affordability", x = "Affordability percentage rank", y = "Low-Use percentile rank")
  SaveFig("LowUseAffordcolour.pdf")

```

##MSOA Data

```{r}
MSOAModelData <-CreateGeogModelData("MSOA", file.path(basewd, "BootstrapMSOA")) 
MSOAMean <- MeanModelData(MSOAModelData, "MSOA") %>%
  left_join(select(MSOAtoLAD, MSOA11CD, LAD11CD)) %>%
  mutate(AffordRank2 = scale(AffordRank)^2,
         AffordRatio2 = scale(AffordRatio)^2,
         TourismDensityRank = rank(TourismDensity, ties.method = "max")) %>%
  left_join(select(LADMean, LAD11CD, LADHighLUP = HighLUP, LADHighVal = HighVal, #Add in Saucy LAD variables
                   LADAffordRatio = AffordRatio, LADAffordRatio2 = AffordRatio2,
                   LADTourismDensity = TourismDensity, LADMedianDiff = MedianDiff), by = "LAD11CD") %>%
  mutate(BothHighLUP =case_when(
    "TRUE"==HighLUP & "TRUE" ==LADHighLUP ~ 1, # all high
    "TRUE" == HighLUP & "FALSE" ==LADHighLUP ~ 2, #High LUP in MSOA but not in LAD
    "FALSE"==HighLUP & "TRUE" ==LADHighLUP ~ 3, #High LUP in LAD but not MSOA
    "FALSE"==HighLUP & "FALSE"==LADHighLUP ~ 4 #LOW LUP in both LAD and MSOA
  ) %>% as.factor,
  BothHighVal =case_when(
    "TRUE" ==HighVal & "TRUE" ==LADHighVal ~ 1, # all high
   "TRUE" == HighVal & "FALSE"==LADHighVal ~ 2, #High LUP in MSOA but not in LAD
    "FALSE"==HighVal & "TRUE" ==LADHighVal ~ 3, #High LUP in LAD but not MSOA
    "FALSE"==HighVal & "FALSE"==LADHighVal ~ 4 #LOW LUP in both LAD and MSOA
  ) %>% as.factor,
  MSOALUPLADVal =case_when(
    "TRUE" ==HighLUP & "TRUE" ==LADHighVal ~ 1, # all high
   "TRUE" == HighLUP & "FALSE"==LADHighVal ~ 2, #High LUP in MSOA but not in LAD
    "FALSE"==HighLUP & "TRUE" ==LADHighVal ~ 3, #High LUP in LAD but not MSOA
    "FALSE"==HighLUP & "FALSE"==LADHighVal ~ 4 #LOW LUP in both LAD and MSOA
  ) %>% as.factor
           )

 MSOAMean %>%
   #filter(TourismDensity!= 0) %>%
  ggplot(aes(x= AffordRank, y = TourismDensity, colour = HighLUP)) + geom_point()

  MSOAMean %>%
   #filter(TourismDensity!= 0) %>%
  ggplot(aes(x= AffordRank, y = TourismDensity, colour = BothHighLUP)) + geom_point()

   MSOAMean %>%
   filter(TourismDensity!= 0) %>%
  ggplot(aes(x= AffordRank, y = TourismDensity, colour = HighVal)) + geom_point()

      MSOAMean %>%
   filter(TourismDensity!= 0) %>%
  ggplot(aes(x= AffordRank, y = TourismDensity, colour = BothHighVal)) + geom_point()

  MSOAMean %>%
  ggplot(aes(x= MedianDiffRank, y = LowUseRank, colour = BothHighVal)) + geom_point()

```

Plots that help explain the data seperation
```{r}
bind_rows(MSOAMean %>% mutate(type = "MSOA"), LADMean %>% mutate(type = "LAD")) %>%
  ggplot(aes(x= AffordRank, y = TourismDensity, colour = HighLUP)) + geom_point() +
  facet_wrap(~type) + 
  labs(title = "Does the geography have above the median LUPs", x = "Affordability Rank", y = "Tourism Density") +
  guides(fill=guide_legend(title="Above Median"))
SaveFig("HighLUP.pdf")

bind_rows(MSOAMean %>% mutate(type = "MSOA"), LADMean %>% mutate(type = "LAD")) %>%
  ggplot(aes(x= AffordRank, y = TourismDensity, colour = HighVal)) + geom_point() +
  facet_wrap(~type) + 
  labs(title = "Is the median LUP price higher than the Median price of all properties", x = "Affordability Rank", y = "Tourism Density") +
  guides(fill=guide_legend(title="Above Median"))
SaveFig("HighVal.pdf")

```

#Affordability

This chunk includes affordability and plots the theoretical and observed curves.
Uses the Bootstrapped values to calculate the affordability

```{r}
MSOAAffordRatio <- MSOAMean %>%
  left_join(select(MSOAtoLAD, MSOA11CD, Region), by ="MSOA11CD")

MSOAAffordRatio  <- bind_rows(MSOAAffordRatio %>% filter(Region != "London") %>% mutate(type = "Excluding London"), 
                    MSOAAffordRatio %>% mutate( type = "Including London"),
                    MSOAAffordRatio %>% filter(Region == "London") %>% mutate( type = "Only London")) %>%
    group_by(type) %>%
    mutate(
      decile = ntile(AffordRatio, 100)) %>%
      group_by(decile, type) %>%
    summarise(mean = mean(LowUsePerc),
              median = median(LowUsePerc),
              counts = n(),
              AffordRatio= mean(AffordRatio)
              )

MSOAAffordRatio %>% 
  #filter(type == "Excluding London") %>% 
  ggplot(aes(x= decile, y = median)) + 
  facet_grid(.~type) + 
  #geom_point(data =unaffordExLon , aes(x= decile, y = LowUsePerc), alpha = 0.2 )+
  geom_point(aes(colour = type))  +
  labs(x = "2 percentile group", y = "median Percent Low Use") + 
  theme(legend.position= "none") + 
  geom_smooth(method="loess",se=FALSE, colour = "black") +
 #coord_cartesian(ylim = c(0.01, 0.05)) +
   scale_y_continuous(labels = scales::percent)
SaveFig("LowUsePercVSratio.pdf")

#The plot of the theoretical curves
data_frame(Index = 1:100) %>%
  mutate(Primary = exp(-0.05*(Index))+0.05, 
         Auxilary = (exp(Index/100)-1),
    Total = Primary + Auxilary) %>%
  gather(key = "Type", value = "value", -Index) %>%
  ggplot(aes(x= Index, y = value, colour = Type )) +
  geom_line() +
  labs(title = "LUP percentage as a function of primary and Auxilary demand",
       x = "Demand",
       y = "LUP percentage") +
  theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank()
  )
SaveFig("TheoreticalDemand.pdf")

```

#Models

The models that will be made are all binary logistic regression

Are the majority of the of the LUPS above the LAD price? 
  this uses the variables Affordability Rank (normalised), the square of the affordability rank and tourism
Are there more than the average LUPS in the LAD

Are there more than average LUPS in the LSOA/MSOA

Both models can be bootstrapped as the investment number relies on the inferred value of Homes and LUPs and the affordability relies on the bootstrapped Homes price... only question is is there enough variation in the median homes price for it to be worth it.

#Model Combinations

The below are the combinations of formula and outcome variable

```{r}

Outcomes <- c("HighVal", "HighLUP")
OutcomesLinear <- c("LowUsePerc", "MedianDiff")

Formulas <- c(as.formula(Reference ~ AffordRatio),
   as.formula(Reference ~ TourismDensity + AffordRatio),
      as.formula(Reference ~ AffordRatio + AffordRatio2),
   as.formula(Reference ~ TourismDensity + AffordRatio + AffordRatio2))

```

#LAD Model

```{r}

cv <- tidy(vfold_cv(mtcars, v = 5))

cv <- tidy(vfold_cv(mtcars, v = 5, repeats = 2))

#Create sample
set.seed(1983)
TestResample <- 1:100 %>% map(~  resample_partition(LADMean , c(test = 0.2, train = 0.8)))

#Create a load of classifiers
LADModels <- LinearClassifierCombiner(LADMean,  Outcomes, Formulas, TestResample) %>%
  mutate(IncAfford2 = grepl("2", Formula),
           IncTourism = grepl("Tourism", Formula)
         )

#Create a load of linear models
LADModelsLin <- LinearClassifierCombiner(LADMean,  OutcomesLinear, Formulas, TestResample, Classifier = FALSE) %>%
  mutate(IncAfford2 = grepl("2", Formula),
           IncTourism = grepl("Tourism", Formula)
         )

LADModels %>%
  group_by(Dependent, IncAfford2, IncTourism) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            ) %>%
  ungroup %>%
   mutate(Dependent = ifelse(Dependent=="HighLUP", "High LUP %", "High LUP value")) %>%
  rename(Afford2 = IncAfford2,
         Tourism = IncTourism) %>% mutate(ModelID = rep(1:4,2))

#Linear models are rubbish
LADModelsLin %>%
  select(-sample) %>%
  group_by(Dependent, IncAfford2, IncTourism) %>%
  summarise_if(is.numeric, funs(mean))

#Explore the best classifier
#There is no structure which adequately models whether A LAD will have more than the median LUP percentage
LADMODOut2 <- ExtractClassCoeffs(LADMean, "HighLUP", TestResample, Formulas[[1]])
LADMODOut2 %>% ggplot( aes(x = term, y = estimate)) + geom_boxplot() #Model Coefficients
LADMODOut2 %>% ggplot( aes(x = term, y = p.value)) + geom_boxplot() #Model p.values

#Formula two provides the most accurate model without over fitting. Being a good model for identifying which LADs will have hhigh value LUPs
LADMODOut <- ExtractClassCoeffs(LADMean, "HighVal", TestResample, Formulas[[2]])
LADMODOut %>% ggplot( aes(x = term, y = estimate)) + geom_boxplot() #Model Coefficients
LADMODOut %>% ggplot( aes(x = term, y = p.value)) + geom_boxplot() #Model p.values

LADMODOut %>%
  select(-statistic, -std.error) %>%
  group_by(term) %>%
  summarise_all(funs(mean, median))

```

#MSOA Model
```{r}
set.seed(1983)
TestResample <- 1:100 %>% map(~  resample_partition(MSOAMean , c(test = 0.2, train = 0.8)))

MSOAModels <-LinearClassifierCombiner(MSOAMean,  Outcomes, Formulas, TestResample) %>%
  mutate(IncAfford2 = grepl("2", Formula),
           IncTourism = grepl("Tourism", Formula)
         )

MSOAModelsLin <- LinearClassifierCombiner(MSOAMean,  OutcomesLinear, Formulas, TestResample, Classifier = FALSE) %>%
  mutate(IncAfford2 = grepl("2", Formula),
           IncTourism = grepl("Tourism", Formula)
         )

MSOAModels %>%
  group_by(Dependent, IncAfford2, IncTourism) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            ) %>%
  ungroup %>%
   mutate(Dependent = ifelse(Dependent=="HighLUP", "High LUP %", "High LUP value")) %>%
  rename(Afford2 = IncAfford2,
         Tourism = IncTourism)

#Including both tourism and the quadratic affordability term produces the best model. Which makes sense given the plot earlier
LADMODOut <- ExtractClassCoeffs(MSOAMean, "HighLUP", TestResample, Formulas[[4]])
LADMODOut %>% ggplot( aes(x = term, y = estimate)) + geom_boxplot() #Model Coefficients
LADMODOut %>% ggplot( aes(x = term, y = p.value)) + geom_boxplot() #Model p.values

#Formula two provides the most accurate model without over fitting. Being a good model for identifying which LADs will have hhigh value LUPs
MSOAMODOut2 <- ExtractClassCoeffs(MSOAMean, "HighVal", TestResample, Formulas[[1]])
MSOAMODOut2 %>% ggplot( aes(x = term, y = estimate)) + geom_boxplot() #Model Coefficients
MSOAMODOut2 %>% ggplot( aes(x = term, y = p.value)) + geom_boxplot() #Model p.values

MSOAMODOut %>%
  select(-statistic, -std.error) %>%
  group_by(term) %>%
  summarise_all(funs(mean, median))

#Rubbish
MSOAModelsLin %>%
  select(-sample) %>%
  group_by(Dependent, IncAfford2, IncTourism) %>%
  summarise_if(is.numeric, mean, na.rm = T)

```

Take away message is that the LAD model cannot predict LUP% whilst the MSOA model cannot predict High LUP Value