### 2018-08-17 MSOAmodel.Rmd
Cleaning and don't ruin postocdes

---
title: "Untitled"
author: "Jonathan Bourne"
date: "31 July 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

Use the MSOA to make a model it has many examples and maps perfectly to the income estimates

Number of MSOA per LAD
```{r}

testMSOA <- EW2 %>%
  group_by(LAD11CD, MSOA11CD) %>%
  summarise(counts = n()) %>%
  group_by(LAD11CD) %>%
  summarise(counts = n())

testWard <- EW2 %>%
  group_by(LAD11CD, Admin_ward_code) %>%
  summarise(counts = n()) %>%
  group_by(LAD11CD) %>%
  summarise(counts = n())

```

#LAD quartiles are wrong

```{r}

setwd("/home/jonno/Dropbox/SSE/Empty Homes/BootStrapAuthorities2")
test <- DistribCompareBootstrapper(ChelseaLONDATA, 1983, 501, type = NULL, PropertyTypes, GroupVars = "class"  )

ChelTest1 <- DistribCompareBootstrapper(ChelseaLONDATA, 1983, 501, type = NULL, PropertyTypes, GroupVars = "MSOA11CD" )

test2 <- DistribCompareBootstrapper(ChelseaLONDATA, 1983, 501, type = NULL, PropertyTypes, GroupVars = c("MSOA11CD", "class"))

MSOAList <- unique(test2$MSOA11CD) %>%
  map(~{
test2 %>%
  filter(MSOA11CD == .x)
})

test3 <- MSOAList[[21]] %>%
  CleanForPlotClassDiff

MSOAList[[21]] %>%
CleanForPlotClassDiff()  %>%
PlotClassDiff(., "Kensington and Chelsea by MSOA")

test %>%
  mutate(LowUse = Homes) %>%
  CleanForPlotClassDiff()  %>%
  PlotClassDiff(., "Kensington and Chelsea full borough")

test %>%
#  mutate(LowUse = Homes) %>%
  CleanForPlotClassDiff2()  %>%
  PlotClassDiff(., "Kensington and Chelsea full borough")

CleanForPlotClassDiff2 <- function(df){ #relative to the homes distrib
  df %>% 
  group_by(ID, class) %>%
  summarise(Homes = sum(Homes), LowUse = sum(LowUse)) %>%
  group_by(ID)%>%
  mutate(RatioExvsAct = LowUse/sum(LowUse),
         HomesRatio = Homes/sum(Homes),
         RatioExvsAct = (RatioExvsAct/HomesRatio)-1)
}

```

```{r}

PropertyTypes <- c("D", "S", "T", "F")
setwd("/home/jonno/Dropbox/SSE/Empty Homes/BootstrapMSOA")

#Find what needs to be bootstrapped
BootStrapFiles <- list.files(getwd()) %>% gsub("DATA.rds", "",.)
  newLADsnames <- sub("DATA", "", ls(pattern = "DATA", envir = globalenv())) %>% 
    .[!(. %in% BootStrapFiles)] 
  newLADs <- paste0(newLADsnames, "DATA")

  test <- ls(pattern = "DATA") %>% map_df(~
 data.frame(LAD = .x, LSOA = get(.x) %>% nrow(),Homes = sum(get(.x)$Homes, na.rm = T))) %>%
    arrange(Homes)

  newLADs <- newLADs[match(test$LAD, newLADs)] %>% .[!is.na(.)]

  #only bootstrap if there is something to add.
  #Save each boostrap as you go along to avoid losing it all in a crash
  if(length(newLADsnames)>0){
    newLADs  %>% walk(~{print(.x)
      DistribCompareBootstrapper(get(.x), 1652, 501, type=NULL, PropertyTypes, GroupVars = "MSOA11CD", Limit =1.3e5  )%>%
      saveRDS(., file = paste0(.x, ".rds"))
      gc()
      }) 
  }

#once all Bootratps complete load them all into a list
 BootStrapRES <- list.files(getwd()) %>%
   map(~readRDS(.x))

 #name the list
 names(BootStrapRES) <- list.files(getwd()) %>% gsub(".rds", "",.)

```

```{r}

library(Rtsne)

# dfmat <- BootDeetsMean %>%
#   select(AffordDiff, Decile, AffordRatio, AffordRatio2, DecileDiff) %>%
#   as.matrix()

dfmat <- ModelMSOAdf %>%
  select( Tourism, AffordRatio, AffordRatio2) %>%
  as.matrix()

set.seed(12358)
TsnecorOut1 <-dfmat  %>%
  Rtsne(verbose = TRUE, eta = 1, max_iter = 5000, perplexity = 50, check_duplicates = FALSE)

# tsne_corplot <- data.frame(x1 = TsnecorOut1$Y[,1],
#                         y1 = TsnecorOut1$Y[,2]) %>%
#   bind_cols(BootDeetsMean) %>%
#   mutate(MedianCheck = MSOAHomesMedian<MSOALowUseMedian)

tsne_corplot <- data.frame(x1 = TsnecorOut1$Y[,1],
                        y1 = TsnecorOut1$Y[,2]) %>%
  bind_cols(ModelMSOAdf)

tsne_corplot %>%
    ggplot() +
  geom_point(aes(x=x1, y=y1, color= TourDepCat==1|TourDepCat==2))

BootDeetsMean %>%
  ggplot(aes(x = LowUseConc, y = AffordDiff, colour = Reference)) + geom_point()

BootDeetsMean %>%
  filter(sales <1) %>%
  ggplot(aes(x = sales, y = LowUse, colour = Reference)) + geom_point()

```

#Sample quantiles

```{r}

#Create a data frame of LSO
PriceCounts <-prices %>%
  filter(X5 %in% c("D", "S", "T", "F")) %>%
  group_by(LSOA11CD) %>%
  summarise(counts = n()) %>%
  arrange(counts)
##The subsampling is 1/10 of the number of homes in each LSOA, this stops crashes
PriceCounts<- EW2 %>%
  select(LSOA11CD = ECODE, Homes, Pop, LAD11CD, MSOA11CD) %>%
  left_join(.,PriceCounts, by = "LSOA11CD") %>%
  mutate(counts = ifelse(is.na(counts), 0, counts),
         Homes2 = Homes/10) %>%
  filter(counts!=0)

#Create a list of price vectors by LSOA
#This process can take a while
setwd(basewd)
if(file.exists("LSOAPriceList2.rds")){

  LSOAPriceList <- readRDS("LSOAPriceList2.rds")

}else{

  LSOAPriceList <- 1:nrow(PriceCounts) %>% map(~{
   print(.x)
    prices %>%
      filter(LSOA11CD==PriceCounts$LSOA11CD[.x], 
             X5 %in% c("D", "S", "T", "F")) %>%
      pull(X2)
    })
  names(LSOAPriceList) <- PriceCounts$LSOA11CD
  saveRDS(LSOAPriceList, "LSOAPriceList2.rds")

}

if(file.exists("AllDataQuartiles.rds")){

  AllDataQuartiles <- readRDS("AllDataQuartiles.rds")

}else{
  #Subset list to only the LADs used in analyis
  LSOAPriceList2 <- LSOAPriceList[names(LSOAPriceList) %in% unique(DATAdf$LSOA11CD)]
  PriceCounts2 <- PriceCounts %>%
    filter(LSOA11CD %in% unique(DATAdf$LSOA11CD))

  AllDataQuartiles <- StratifiedBoot(LSOAPriceList2, PriceCounts2$LSOA11CD, PriceCounts2$Homes2, samples = 501)

  saveRDS(AllDataQuartiles, "AllDataQuartiles.rds")

  rm(list = c("LSOAPriceList2", "PriceCounts2"))

}

rm(LSOAPriceList)

AllDataQuartilesSummary <- AllDataQuartiles %>%
  summarise_all(funs(mean, sd))

BasicQuartiles <- prices$X2 %>% .[prices$LAD11CD %in% unique(DATAdf$LAD11CD)] %>% quantile(.)

#difference between the two up to 8% this has knock on effects later in the analysis if the quartiles are not calculated correctly
AllDataQuartilesSummary[2:4]/BasicQuartiles[2:4]

prices <- prices %>%
  mutate(CountryClass = cut(X2, AllDataQuartilesSummary[2:4] %>% c(0,., Inf), 
                              labels =     c("Lower", "Lower-Mid", "Upper-Mid", "Upper"), 
                              right = F) %>% fct_relevel(., "Upper", after = 3))

rm(BasicQuartiles);rm(AllDataQuartiles);rm(AllDataQuartilesSummary)

```

#Calc Real Quantiles for all data

This takes ages and gives an almost identical number to simple running quantiles on the whole dataset

```{r Real Quantile Chunk}

test <-  AllDataQuartiles %>%
   select(-X0.,-X100.) %>%
   summarise_all(funs(mean, sd)) %>%
   mutate(X25Perc = X25._sd/X25._mean,
          X50Perc = X50._sd/X50._mean,
          X75Perc = X75._sd/X75._mean)

prices %>%
   filter(X5 %in% c("D", "S", "T", "F")) %>%
  pull(X2) %>% quantile()

```

#MSOA Quartiles

I'm not sure this is useful

```{r}
#Create a list of vectors where each element of the list is an MSOA and the vector elements are its LSOA
 LSOAinMSOA <- unique(PriceCounts$MSOA11CD) %>% map(~{
   print(.x)
    PriceCounts %>%
      filter(MSOA11CD==.x) %>%
      pull(LSOA11CD)
 })
 #name the elements of the list
 names(LSOAinMSOA) <- unique(PriceCounts$MSOA11CD)

 #Create a bootstrap for each of the MSOA

 MSOABootQuartiles <- unique(PriceCounts$MSOA11CD) %>% map_df(~{
   print(.x)
   PriceCounts2 <- PriceCounts %>%
     filter(MSOA11CD == .x)

   LSOAPriceList2 <- LSOAPriceList[names(LSOAPriceList) %in% PriceCounts2$LSOA11CD]

   StratifiedBoot(LSOAPriceList2, PriceCounts2$LSOA11CD, PriceCounts2$Homes2, samples = 100) %>%
     mutate(MSOA11CD=.x)

 })

 saveRDS(MSOABootQuartiles, "MSOABootQuartiles.rds")

 MSOABootQuartiles2 <- MSOABootQuartiles %>%
   select(-X0.,-X100.) %>%
   group_by(MSOA11CD) %>%
   summarise_all(funs(mean, sd)) %>%
   mutate(X25Perc = X25._sd/X25._mean,
          X50Perc = X50._sd/X50._mean,
          X75Perc = X75._sd/X75._mean)  %>%
   left_join(., EW2%>%
               select(MSOA11CD, LAD11CD, LAD11NM) %>% distinct()) %>%
   left_join(., IncomeEst) %>%
   select(MedianPrice=X50._mean, Yearly.income, MSOA11CD, LAD11CD, LAD11NM) %>%
   mutate(AffordRatio = MedianPrice/Yearly.income,
          AffordRank = percent_rank(AffordRatio))
```

#Local value

Here I create a graph of the MSOA. 
I calculate the avereage distance by LAD.
I then use that distance to find all the MSOA within that jump distance.
I then calculate the mean and quartiles of the local area for each MSOA
From there I can calculate the local area affordability and thus the relative local area affordability

```{r}

 MSOAAdjacency <- readShapeSpatial(file.path("/home/jonno/Dropbox/SSE/Empty Homes/ShapeFiles/Middle_Layer_Super_Output_Areas_December_2011_Super_Generalised_Clipped_Boundaries_in_England_and_Wales",
                                             "Middle_Layer_Super_Output_Areas_December_2011_Super_Generalised_Clipped_Boundaries_in_England_and_Wales.shp"))

 test <- poly2nb(MSOAAdjacency)

 MSOAAdjMat = nb2mat(test, style="B",zero.policy=T)

#Add MSOA codes so that they are added as the vertex names
colnames(MSOAAdjMat) <- MSOAAdjacency$msoa11cd

MSOAgraph <- graph.adjacency(MSOAAdjMat, mode = "undirected")

#Add in Lad

MSOAgraph <- set_vertex_attr(MSOAgraph, 
                name = "LAD11CD", 
                value = MSOAtoLAD$LAD11CD[match(get.vertex.attribute(MSOAgraph, "name"),
                                          MSOAtoLAD$MSOA11CD)]
                )

#Subset by LAD into a list of graphs

Subgraph <- function(g, KeepNodes){

  VertexNames <-get.vertex.attribute(g, "name")
  VertexNames<- VertexNames[!(VertexNames %in% KeepNodes)]

g2 <- delete_vertices(g, VertexNames)

g2
}

AllLADsGraph <- unique(MSOAtoLAD$LAD11CD) %>% map(~{
  KeepNodes <-MSOAtoLAD %>%
  filter(LAD11CD ==.x)
KeepNodes2 <- KeepNodes$MSOA11CD

  Subgraph(MSOAgraph, KeepNodes2)
})

#name the elements by the LAD represented
names(AllLADsGraph) <- unique(MSOAtoLAD$LAD11CD)

#calculate mean distance and diameter of each lad
MSOAconnections <- AllLADsGraph %>% map_df(~{

  tibble(distance = mean_distance(.x), diameter = diameter(.x), MSOA = vcount(.x))

}) %>%
  mutate(LAD11CD = names(AllLADsGraph)) %>%
  filter(MSOA !=1) #removes city of london which only has a single MSOA and so can't be included in the calc

MSOAconnections %>%
  ggplot(aes(x=diameter )) + geom_density()

#Using mean distance is best
summary(MSOAconnections$distance)
summary(MSOAconnections$diameter)

# find names of local MSOA using an x jump distance
EgoSize <- tibble(ego_size = ego_size(MSOAgraph, 2), MSOA11CD = get.vertex.attribute(MSOAgraph, "name")) %>%
  left_join(MSOAtoLAD)
##Ego size of jump two has a mean almost identical to the mean number of MSOA in a LAD. However the Ego size is less skewed.
mean(EgoSize$ego_size)

test <-  tibble(MSOA11CD = ego(MSOAgraph, 1, nodes = "E02000001")[[1]] %>% names) %>%
  left_join(MSOAtoLAD)

##names of the MSOA in each area.
MSOAEgoList <- ego(MSOAgraph, 2) %>%map(~names(.x))
names(MSOAEgoList) <- get.vertex.attribute(MSOAgraph, "name")

```

```{r}
#Create a list of vectors where each element of the list is an MSOA and the vector elements are its LSOA
 LSOAinMSOAEgo <- MSOAEgoList %>% map(~{
   #print(.x)
    PriceCounts %>%
      filter(MSOA11CD %in% .x) %>%
      pull(LSOA11CD)
 })

MSOAEgoList1 <- ego(MSOAgraph, 1) %>%map(~names(.x))
names(MSOAEgoList1) <- get.vertex.attribute(MSOAgraph, "name")

MSOAEgoList2 <- ego(MSOAgraph, 2) %>%map(~names(.x))
names(MSOAEgoList2) <- get.vertex.attribute(MSOAgraph, "name")

 TourismMSOAEgo2 <- MSOAEgoList %>% map_dbl(~{
   print(.x)
    Tourism %>%
      filter(MSOA11CD %in% .x) %>%
      pull(Tourism) %>% sum
 }) %>%
   tibble(Ego2Tourism = .,
          MSOA = names(MSOAEgoList))

  TourismMSOAEgo1 <- MSOAEgoList1 %>% map_dbl(~{
   print(.x)
    Tourism %>%
      filter(MSOA11CD %in% .x) %>%
      pull(Tourism) %>% sum
 }) %>%
   tibble(Ego1Tourism = .,
          MSOA = names(MSOAEgoList1))

 ##name the elements of the list
 names(LSOAinMSOAEgo) <- names(MSOAEgoList)

 ##Create a bootstrap for the ego of each MSOA
 NeighbourhoodQuartiles <- names(MSOAEgoList) %>% map_df(~{

   print(.x)
   target <- MSOAEgoList[[.x]]
   PriceCounts2 <- PriceCounts %>%
     filter(MSOA11CD %in% target)

   LSOAPriceList2 <- LSOAPriceList[names(LSOAPriceList) %in% PriceCounts2$LSOA11CD]

   Out <-StratifiedBoot(LSOAPriceList2, PriceCounts2$LSOA11CD, PriceCounts2$Homes2, samples = 10) %>%
     mutate(MSOA11CD=.x)

   Out

 })

```

Do westminster and lambeth fall into each others local area?
need to check and make sure they do!
```{r}
#Calculate mean local income
LocalIncome <- names(MSOAEgoList) %>% map(~{
  print(.x)
#.x <- "E02000001"
Group <-MSOAEgoList[[.x]]

Incomest2 <- IncomeEst %>%
  left_join(select(MSOAtoLAD, MSOA11CD, Homes),  by="MSOA11CD")

Out <- Incomest2 %>%
  filter(MSOA11CD %in% .x) %>%
  summarise(LocalIncome =sum(Yearly.income* Homes)/sum(Homes)) %>% pull()

Out
}) %>% unlist %>%
  tibble(MSOA11CD = names(MSOAEgoList), LocalIncome = .)

LocalAfford <-NeighbourhoodQuartiles %>%
  select(-ID,-X0., -X100.) %>%
    group_by(MSOA11CD) %>%
  summarise_all(funs(mean, sd, min, max)) %>%
  mutate(perc = X50._sd/X50._mean) %>%
  left_join(LocalIncome, by = "MSOA11CD") %>%
  mutate(LocalAfford = X50._mean/LocalIncome) %>%
  select(MSOA11CD, MedianLocalHomesPrice = X50._mean, LocalAfford)

```

#Complex models

These models work at MSOA level only and include data from the LAD level
They don't really add anything extra so aren't included in the main work. Just kept here so I don't forget I have already checked.

```{r}
MSOAMean2 <-MSOAMean %>%
  mutate(RelativeAfford = AffordRatio/LADAffordRatio,
         RelativeTourismDens = TourismDensity/LADTourismDensity,
         RelativeTourismDens = ifelse(is.finite(RelativeTourismDens), RelativeTourismDens, 100),
         HighFinance = factor("TRUE"==LADHighVal & "TRUE"==HighLUP, levels = c("TRUE", "FALSE"))) %>%
  filter(!is.na(HighVal)) #Temporary  until I work out how to get rid of those weird MSOA

KitchenSink <- as.formula(Reference ~ TourismDensity+ AffordRatio+ AffordRatio2+ RelativeAfford +RelativeTourismDens+
                            LADTourismDensity+ LADAffordRatio+ LADAffordRatio2+ LADMeanMedianRatio)

KitchenSink2 <- as.formula(Reference ~RelativeAfford + RelativeTourismDens +
                            LADTourismDensity + AffordRatioScale + AffordRatioScale2 + LADMeanMedianRatio)

Formulas2 <- c(as.formula(Reference ~ TourismDensity+ AffordRatio+ AffordRatio2),
               KitchenSink)

####Needs updating
MSOAModels2 <-LinearClassifierCombiner(MSOAMean2,  Outcomes, Formulas2, TestResample) %>%
  mutate(IncAfford2 = grepl("2", Formula),
           LAD = grepl("LAD", Formula)
         )

#No difference
MSOAModels2 %>%
  group_by(Dependent, IncAfford2, LAD) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            ) %>%
  ungroup %>%
   mutate(Dependent = ifelse(Dependent=="HighLUP", "High LUP %", "High LUP value"))

#Just under 76%. Better than all other models...but is it?
#The more LAD variables that are included the better the model gets... Until you find out that this is becuase the model uses the lad data to predict the results of other MSOA in the same lad.
ResOut <- MSOAMean2 %>%
  mutate(Reference = HighLUP) %>% 
  TreeResampledModelPerf(., MSOAHighLUPResample, KitchenSink2)

#Data contains LAD information I account for that here

ResOut2 <- MSOAMean2 %>%
  mutate(Reference = HighLUP) %>%
  TreeResampledModelPerf(., LADHighLUPResample, KitchenSink2, LADData = LADMean )

#The model was getting information from the shared LAD data so making it appear much better than it really is
ResOut %>% SummariseModels()
ResOut2 %>% SummariseModels()

```

```{r}
#This model uses information from the over all LUP as a result I need to subset the data by lups then train on the MSOA
set.seed(1983)
TestResample <- vfold_cv(LADMean, v = 5, repeats = 20)

ComplexModels <- c("BothHighLUP", "BothHighVal", "MSOALUPLADVal", "LADHighLUP", "HighLUP", "LADHighVal") %>%
  map_df(~{
    print(.x)
    ResOut2 <- MSOAMean2 %>%
      rename_(Reference = .x) %>%
      TreeResampledModelPerf(., TestResample, KitchenSink2, LADData = LADMean) %>%
      mutate(Dependent = .x)
})

#the data from the Lup is increasing the accuracy of the model
ResOut2 <- MSOAMean2 %>%
  mutate(Reference = HighLUP) %>%
  TreeResampledModelPerf(., TestResample, KitchenSink2, LADData = LADMean)

ComplexModels %>%
  group_by(Dependent) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )

KitchenSink2 <- as.formula(Reference ~RelativeAfford +RelativeTourismDens+
                            LADTourismDensity+ LADAffordRatio)

ComplexModels2 <- c("BothHighLUP", "BothHighVal", "MSOALUPLADVal", "LADHighLUP", "HighLUP", "LADHighVal") %>%
  map_df(~{
    print(.x)
    ResOut2 <- MSOAMean2 %>%
      rename_(Reference = .x) %>%
      TreeResampledModelPerf(., TestResample, KitchenSink2, LADData = LADMean) %>%
      mutate(Dependent = .x)
})

ComplexModels2 %>%
  group_by(Dependent) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )

```

Highly financialized

```{r}
KitchenSink2 <- as.formula(Reference ~RelativeAfford + RelativeTourismDens +
                            LADTourismDensity + AffordRatioScale + AffordRatioScale2 + LADMeanMedianRatio, Ego2TourismDens2+Ego2TourismDens)

test <- MSOAMean2 %>%
  select(HighFinance, LADHighVal, HighLUP)

ComplexModels2 <- c("HighFinance", "HighLUP", "LADHighVal") %>%
  map_df(~{
    print(.x)
    ResOut2 <- MSOAMean2 %>%
      rename_(Reference = .x) %>%
      TreeResampledModelPerf(., TestResample, KitchenSink2, LADData = LADMean) %>%
      mutate(Dependent = .x)
})

ComplexModels2 %>%
  group_by(Dependent) %>%
  summarise('Beats NULL' = sum(Accuracy>AccuracyNull)/n(), #How many times did the models beat NULL
            Accuracy = mean(Accuracy), #What was the mean accuracy
            'NULL Model' = mean(AccuracyNull),
           'Null Difference' = mean(Accuracy-AccuracyNull) #Mean difference relative to NULL
            )
```

#Collaspy tree voting
```{r}

test <- MSOAMean2 %>% left_join(MSOAtoLAD) %>%
      rename_(Reference = "LADHighLUP") %>%
MSOATreeVoting(., TestResample, KitchenSink2 )

test %>% SummariseModels()

```